### 2015-10-26 report/paper.tex

\documentclass{article} % For LaTeX2e
% We will use NIPS submission format
\usepackage{nips13submit_e,times}
% for hyperlinks
\usepackage{hyperref}
\usepackage{url}
% For figures
\usepackage{graphicx} 
\usepackage{subfigure} 
% math packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsopn}
\usepackage{ifthen}
\usepackage{natbib}

\title{Project-I by Group CITYNAME}

\author{
Corentin Tallec\\
Roman Shirochenko\\
EPFL \\
\texttt{corentin.tallec@epfl.ch\\
roman.shirochenko@epfl.ch} \\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\nipsfinalcopy

\begin{document}

\maketitle

\begin{abstract}
	In this report we summarize the results we obtained for the first machine
	learning of the EPFL machine learning course. Facing two datasets, we fit
	linear models and perform linear regression on the first one, and perform data
	classification using logistic regression on the second one. For the first
	dataset, we find that simple linear regression and ridge regression gives
	decent results, but that we can get better results by isolating one simple
	feature and performing transformations on that feature.
\end{abstract}

\section{Linear Regression}
\subsection{Data Description}
Our first data set consisted of a train set, with an input and an output
variable respectively $\mathbf{X}$ and $\mathbf{y}$, composed of $N=2800$ data
samples, and a test set with only $\mathbf{X}$ variable observed and $N=1200$
data samples. $\mathbf{X}$ spans over $D=77$ features, among which 9 of them are
categorical, with number of categories ranging from 2 to 4, 2 of them are binary
and the other are real-valued.

Our goal is to produce linear predictions for test examples and approximate the
test error.

\subsection{Data visualization and cleaning}
We performed basic exploratory data analysis on our data. Figure \ref{fig:boxplotX} shows distributions of all 10 input variables. As expected the data is not centered, and therefore we normalize the data. Figure \ref{fig:histY} shows the histogram of the output variable. We see that there are two data points which have rather high value of the output variables compared to rest of the data. Since these are only two data points, we remove them from the data (example number 362 and 402).

We also investigated the correlation between output and input variables (plot not shown, since it is not important for our analysis). We found that most of the variables are correlated with the output, confirming the need for a full regression analysis. The third variable did not seem significantly correlated, however this preliminary observation does not suggest removal of this variable from the dataset.

We use a dummy encoding for the categorical variable. The binary variable does not require any dummy encoding. This gives us a total of 14 input variables.

The input matrix $\mathbf{X}$ is rank-deficient with a rank of 6, instead of 10. This implies that we must ``lift" the eigenvalues, using methods such ridge regression.

\begin{figure}[!t]
\center
\subfigure[Boxplot of real-valued $\mathbf{X}$. Data is not centered and therefore we normalize it.]{\includegraphics[width=2.5in]{figures/boxplotX.pdf} \label{fig:boxplotX}}
\hfill
\subfigure[Histogram of $\mathbf{y}$. We can clearly see two outliers.]{\includegraphics[width=2.5in]{figures/histY.pdf} \label{fig:histY}}
\caption{}
\end{figure}

\subsection{Ridge regression}
We applied least-squares and ridge regression to this dataset. Since the matrix is ill-conditioned, least-squares is not suitable. Therefore, we report results obtained with ridge regression alone. Note that the improvements using ridge regression were modest and not much lower than that of linear regression, however we do not expect least-squares to work well when there is a lot of testing data available.

Figure \ref{fig:ridgeCurve} shows the results obtained with ridge regression when we use $50\%$ of the data as test data and rest as training data. We varied the value of $\lambda$ from $10^{-4}$ to $10^3$, choosing total 500 points in between. We can see that there is a small improvement obtained for some values of $\lambda$.

We did experiments to plot a learning curve for this data (see Andrew Ng's notes about the learning curve). We held out 20\% of data as test data and rest as training data. We chose to slowly increase the proportion of data used for training. For each proportion of the training data, we repeated the experiment 30 times to compute the distribution of error. We fit ridge regression to each sampled training set and test it on the same $20\%$ test data. We varied the value of $\lambda$ from $10^{-4}$ to $10^3$, choosing total 500 points in between.

This gives us the learning curve shown in Fig. \ref{fig:learnCurve}. The blue curve shows the train error while red curve shows the test error. We can see that both training and test error converge, with the variance of estimates decreasing as we increase the training data size. There is also a very small gap between the train and test errors, showing that the linear model is a reasonable choice. The small gap exists perhaps because we have only limited test data.

\subsection{Feature transformations}
We tried several feature transformations. We found that we get a small improvement in performance when we take $\sqrt{|X_{ni}|}$ for all entries of $\mathbf{X}$. We did not check (due to lack of time) whether it matters if we apply this to one variable or all. We performed experiments similar to the last section (although one should really do cross-validation). Values of lambda were kept same as the last section.

We compare three methods. First is a baseline where we do not use any input variables i.e. mean value of the output. The second method is the ridge regression described in previous section. The third method is ridge regression with a feature transformation. The first method gave RMSE of around 3 which was way worse than the other two methods.

The RMSE for the last two methods are shown in Fig. \ref{fig:summary}. We see that both test and train error decrease, however it appears that the improvement is very little and may not be significant.

\begin{figure}[!h]
\center
\subfigure[Ridge regression for a 50-50 split.]{\includegraphics[width=2.5in]{figures/ridgeCurve.pdf} \label{fig:ridgeCurve}}
\hfill
\subfigure[Comparison of ridge regression with and without feature transformation. The improvement is very little and might be insignificant.]{\includegraphics[width=2.7in]{figures/summary.pdf} \label{fig:summary}}
\subfigure[Learning curve. Blue is training data and red is test data.]{\includegraphics[width=4in]{figures/learningCurve.pdf} \label{fig:learnCurve}}
\caption{}
\end{figure}

\subsection{Summary}
In this report, we analyzed a regression dataset and found that ridge regression is a reasonable fit. We estimate that the average test error is 1.213 ($\pm$ 0.02). We tried some feature transformation and found that there is a small improvement giving us a test error of around 1.198 ($\pm$ 0.015). This improvement, however, is not significant.

\subsubsection*{Acknowledgments}
We would like to thank Timur for making the dataset for this study, Carlos and Ilija to help conduct the experiments, and all the students who attended the session making it exciting for us. All the code, as well as this report, was written by Emti.

\subsubsection*{References}

\end{document}

### 2015-10-27 report/paper.tex
First part of the report written and compiled.

@@
 EPFL \\
-\texttt{corentin.tallec@epfl.ch\\
-roman.shirochenko@epfl.ch} \\
+\texttt{corentin.tallec@epfl.ch} \\
 }
@@
 \subsection{Data visualization and cleaning}
-We performed basic exploratory data analysis on our data.
-Figure \ref{fig:boxplotX} shows distributions of all 10 input variables.
-As expected the data is not centered, and therefore we normalize the data.
-Figure \ref{fig:histY} shows the histogram of the output variable.
-We see that there are two data points which have rather high value of the output variables compared to rest of the data.
-Since these are only two data points, we remove them from the data (example number 362 and 402).
+The first tasks we performed on the data was to visualize it. The first thing we
+did was to obtain a visualization of the data, showed in \ref{fig:boxplotX} and
+\ref{fig:histY}. Those two figures led us to normalize both the input and the
+output variable (normalizing the output variable allow us to have results that are
+easier to interpret, as they are to be compared to values close to 1). No
+obvious outliers are noticeable on the $\mathbf{y}$ histogram.
 
-We also investigated the correlation between output and input variables (plot not shown, since it is not important for our analysis).
-We found that most of the variables are correlated with the output, confirming the need for a full regression analysis.
-The third variable did not seem significantly correlated, however this preliminary observation does not suggest removal of this variable from the dataset.
+We performed featurewise analysis to notice that most of our real valued
+variables have nearly gaussian distributions (or at least unimodal
+distributions), which is not the case of our output variable. Figure
+\ref{fig:cor1} shows how the output variable evolves as a function of the first
+feature. This behaviour is what we notice on most variables, and thus shows no
+clear correlations.
 
-We use a dummy encoding for the categorical variable. The binary variable does not require any dummy encoding. This gives us a total of 14 input variables.
+Although for variables 49 and 66 we notice a multimodal distribution that is
+closer to the behaviour of our output variable, as seen in Figure
+\ref{fig:xhist49} and Figure \ref{fig:xhist66}. We thus chose to try and predict
+$\mathbf{y}$ using only those two features, transformed through polynomial
+transformation.
 
-The input matrix $\mathbf{X}$ is rank-deficient with a rank of 6, instead of 10. This implies that we must ``lift" the eigenvalues, using methods such ridge regression.
+The categorical variables were encoded using dummy encoding, what led to a total
+feature number of 100. As the rank of our dataset is below 100 (it is 70),
+simple linear regression is likely to fail, and we use ridge regression to cope
+for the rank deficiency.
 
@@
 \center
-\subfigure[Boxplot of real-valued $\mathbf{X}$. Data is not centered and therefore we normalize it.]{\includegraphics[width=2.5in]{figures/boxplotX.pdf} \label{fig:boxplotX}}
+\subfigure[Boxplot of $\mathbf{X}$.]{\includegraphics[width=2.5in]{figures/boxplot.jpg} \label{fig:boxplotX}}
+\subfigure[$\mathbf{y}$ as a function of the first feature of $\mathbf{X}$.
+The correlation between the two variables is not
+clear.]{\includegraphics[width=2.5in]{figures/cor1.jpg}\label{fig:cor1}}
 \hfill
-\subfigure[Histogram of $\mathbf{y}$. We can clearly see two outliers.]{\includegraphics[width=2.5in]{figures/histY.pdf} \label{fig:histY}}
+\subfigure[Histogram of
+$\mathbf{y}$. The output distribution is bimodal.]{\includegraphics[width=1.5in]{figures/yhist.jpg} \label{fig:histY}}
+\hspace*{0.3in}
+\subfigure[Histogram of feature 49 of $\mathbf{X}$. We see that the distribution
+	is bimodal, as the output.]{\includegraphics[width=1.5in]{figures/xhist49.jpg}
+\label{fig:xhist49}}
+\hspace*{0.3in}
+\subfigure[Histogram of feature 66 of $\mathbf{X}$. Again, the distribution is
+	bimodal.]{\includegraphics[width=1.5in]{figures/xhist66.jpg}
+\label{fig:xhist66}}
 \caption{}
@@
 \subsection{Ridge regression}
-We applied least-squares and ridge regression to this dataset.
-Since the matrix is ill-conditioned, least-squares is not suitable.
-Therefore, we report results obtained with ridge regression alone.
-Note that the improvements using ridge regression were modest and not much lower than that of linear regression, however we do not expect least-squares to work well when there is a lot of testing data available.
+As the dimension of the data is inferior to the number of feature, we expect
+least-squares to work quite poorly, and thus decided to apply mostly ridge
+regression.
 
-Figure \ref{fig:ridgeCurve} shows the results obtained with ridge regression when we use $50\%$ of the data as test data and rest as training data.
-We varied the value of $\lambda$ from $10^{-4}$ to $10^3$, choosing total 500 points in between.
-We can see that there is a small improvement obtained for some values of $\lambda$.
+For our first attempt at predicting the data, we simply used the initial
+features and used kFold with $k=4$ to display the evolution of the RMSE as a
+function of the regularizing lambda both for the training and test data. 
+The lambda we used ranged from $10^{-4}$ to
+$10^3$, with 500 points. The results of the ridge regression are already
+satisfactory, with an error of approximately $0.105\pm0.001$ on the training set
+and $0.115\pm0.001$ on the test set, that is approximately one tenth of the
+variance of the output variable.
 
-We did experiments to plot a learning curve for this data (see Andrew Ng's notes about the learning curve).
-We held out 20\% of data as test data and rest as training data.
-We chose to slowly increase the proportion of data used for training.
-For each proportion of the training data, we repeated the experiment 30 times to compute the distribution of error.
-We fit ridge regression to each sampled training set and test it on the same $20\%$ test data.
-We varied the value of $\lambda$ from $10^{-4}$ to $10^3$, choosing total 500 points in between.
+We see that for sufficiently low lambda, both the training and test errors do
+not vary much (there is indeed only a very slight variation of test RMSE, with a
+nearly unnoticeable minimum between $\lambda=10$ and $\lambda=100$).
 
-This gives us the learning curve shown in Fig.
-\ref{fig:learnCurve}.
-The blue curve shows the train error while red curve shows the test error.
-We can see that both training and test error converge, with the variance of estimates decreasing as we increase the training data size.
-There is also a very small gap between the train and test errors, showing that the linear model is a reasonable choice.
-The small gap exists perhaps because we have only limited test data.
+The results described here are summerized in the Figure \ref{fig:ridgeCurve}.
 
 \subsection{Feature transformations}
-We tried several feature transformations.
-We found that we get a small improvement in performance when we take $\sqrt{|X_{ni}|}$ for all entries of $\mathbf{X}$.
-We did not check (due to lack of time) whether it matters if we apply this to one variable or all.
-We performed experiments similar to the last section (although one should really do cross-validation).
-Values of lambda were kept same as the last section.
+As notified in a previous section, among all the feature at our disposal,
+features 49 and 66 show distributions that ressemble the ditribution of the
+output variable.
 
-We compare three methods.
-First is a baseline where we do not use any input variables i.e.
-mean value of the output.
-The second method is the ridge regression described in previous section.
-The third method is ridge regression with a feature transformation.
-The first method gave RMSE of around 3 which was way worse than the other two methods.
+This led us to try and regress only on thos two features and polynomial
+transformations of them. Using only those two features, we get an RMSE error
+that is quite close to the one we notified when all the features are used. When
+using polynomial features (up to degree 10), we basically get half the error
+that we got using all the features. Those results were obtained using ridge
+regression. As in the previous section, we obtained the RMSE as a function of
+lambda, calculating both training and test errors using a kFold procedure with
+$k=4$. The results are summarized in Figure \ref{fig:polyridgeCurve}, where, as
+previously mentionned, we used as features only the two features and their
+polynomial transformation with degree ranging from 1 to 10.
 
-The RMSE for the last two methods are shown in Fig.
-\ref{fig:summary}.
-We see that both test and train error decrease, however it appears that the improvement is very little and may not be significant.
+This time, the test error varies much more as a function of lambda, and, given
+our curve, the best lambda possible seems to be around $10^{-6}$.
+
+Finally to evaluate our performances more accurately, we plotted learning curves
+for our model, with 20 features obtained by polynomial transformations and a
+lambda of $10^{-6}$. The learning curve is obtained by using $20\%$ of the
+training data to perform tests, while learning the model on a progressively
+increasing portion of the $80\%$ remaining. We monitor the test and training
+error as functions of the proportion of data used to train. The results are
+summarized in Figure \ref{fig:learningCurve}.
+
+As expected, both training error and test error decrease as the proportion of
+training data increases. The test error obtained seems to exhibit a quit
+satisfactory behaviour.
 
@@
 \center
-\subfigure[Ridge regression for a 50-50 split.]{\includegraphics[width=2.5in]{figures/ridgeCurve.pdf} \label{fig:ridgeCurve}}
+\subfigure[Ridge regression on kFold with
+$k=4$.]{\includegraphics[width=2.5in]{figures/traintestRMSE.jpg} \label{fig:ridgeCurve}}
 \hfill
-\subfigure[Comparison of ridge regression with and without feature transformation.
-The improvement is very little and might be insignificant.]{\includegraphics[width=2.7in]{figures/summary.pdf} \label{fig:summary}}
-\subfigure[Learning curve. Blue is training data and red is test data.]{\includegraphics[width=4in]{figures/learningCurve.pdf} \label{fig:learnCurve}}
+\subfigure[Ridge regression on kFold with feature transformation and
+	$k=4$.]{\includegraphics[width=2.5in]{figures/traintestRMSEpoly.jpg}
+\label{fig:polyridgeCurve}}
+\subfigure[Learning curve. Blue is training data and red is test
+data.]{\includegraphics[width=4in]{figures/learningcurves.jpg}
+\label{fig:learningCurve}}
 \caption{}
@@
 
-\subsection{Summary}
-In this report, we analyzed a regression dataset and found that ridge regression is a reasonable fit.
-We estimate that the average test error is 1.213 ($\pm$ 0.02).
-We tried some feature transformation and found that there is a small improvement giving us a test error of around 1.198 ($\pm$ 0.015).
-This improvement, however, is not significant.
-
-\subsubsection*{Acknowledgments}
-We would like to thank Timur for making the dataset for this study, Carlos and Ilija to help conduct the experiments, and all the students who attended the session making it exciting for us.
-All the code, as well as this report, was written by Emti.
-
-\subsubsection*{References}
-
 \end{document}

### 2015-10-28 Roman code/binLoss.m
Added code for classification part

@@ -1,12 +1,12 @@
 function bin_loss = binLoss(y, y_pred)
 % 1/0 loss error calculation
-    n = length(y);
+    n = size(y_pred,1);
     I = zeros(n,1);
     for i = 1:n
         if y(i) ~= y_pred(i)
             I(i,1) = 1;  
         end
     end
-    bin_loss = 1/n.* sum(I);
+    bin_loss = 1/n * sum(I);
 end
 

### 2015-10-28 Roman code/costFunctionNg.m

@@ -1,0 +1,32 @@
+function [J, grad] = costFunction(theta, X, y)
+%COSTFUNCTION Compute cost and gradient for logistic regression
+%   J = COSTFUNCTION(theta, X, y) computes the cost of using theta as the
+%   parameter for logistic regression and the gradient of the cost
+%   w.r.t. to the parameters.
+
+% Initialize some useful values
+m = length(y); % number of training examples
+
+% You need to return the following variables correctly 
+J = 0;
+grad = zeros(size(theta));
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Compute the cost of a particular choice of theta.
+%               You should set J to the cost.
+%               Compute the partial derivatives and set grad to the partial
+%               derivatives of the cost w.r.t. each parameter in theta
+%
+% Note: grad should have the same dimensions as theta
+%
+
+J = (1/m) * sum(-y.*log(sigmoid(X*theta))-(1-y).*log((1 - sigmoid(X*theta))));
+grad = 1./m * X' *(sigmoid(X*theta) - y);
+
+
+
+
+
+% =============================================================
+
+end

### 2015-10-28 Roman code/costFunctionRegNg.m

@@ -1,0 +1,30 @@
+function [J, grad] = costFunctionReg(theta, X, y, lambda)
+%COSTFUNCTIONREG Compute cost and gradient for logistic regression with regularization
+%   J = COSTFUNCTIONREG(theta, X, y, lambda) computes the cost of using
+%   theta as the parameter for regularized logistic regression and the
+%   gradient of the cost w.r.t. to the parameters. 
+
+% Initialize some useful values
+m = length(y); % number of training examples
+
+% You need to return the following variables correctly 
+J = 0;
+grad = zeros(size(theta));
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Compute the cost of a particular choice of theta.
+%               You should set J to the cost.
+%               Compute the partial derivatives and set grad to the partial
+%               derivatives of the cost w.r.t. each parameter in theta
+
+thetaT = theta;
+thetaT(1) = 0;
+J = (1/m) * sum(-y.*log(sigmoid(X*theta))-(1-y).*log((1 - sigmoid(X*theta)))) + (lambda/(2*m))*(thetaT'*thetaT);
+grad = 1/m * X' *(sigmoid(X*theta) - y) + (lambda/m)* thetaT;
+
+
+
+
+% =============================================================
+
+end

### 2015-10-28 Roman code/dataAnalysis.m

@@ -1,52 +1,25 @@
-% Load data
-clear all;
-load('Rome_regression.mat');
+data = load('Rome_classification');
 
-% plot data histogram
-%figure(1);
-%hist(y_train,100);
-%title('y train')
+tX = data.X_train;
+tX = normalize(tX);
 
-% plot all X dimensions
-%figure(2);
-%for i = 1:77
-%    subplot(10,8,i);
-%    hist(X_train(:,i),100);
-%end
-    
+tY = data.y_train;
 
-%figure(3);
-%boxplot(X_train(:,:));
+alpha = 0.07;
 
-%% Classified data
-% Binary data [0,1]: idx = 9,39,58,61
-idx_zero = [9 39 58 61];
-% Classes [0,2] : idx = 73
-% Classes [1,2] : idx = 3, 47
-% Classes [0, 2, 1] : idx = 2, 57, 67
-% Classes [0, 1, 2, 3] : idx = 7, 12
-idx_classes = [73 3 47 2 57 67 7 12];
-%% 
-% Clean classified data
-idx_X_train = 1:size(X_train,2);
-idx_X_train([idx_zero idx_classes]) = [];
-X_train_cleaned = X_train(:,idx_X_train);
+beta = logisticRegression(tY,tX,alpha);
 
-% Find outliers
-idxOutliers = [];
-for i = 1:size(X_train,2)
-    if i ~= 66
-       continue;
-    end
-    x = X_train(:,i);
-    meanX = mean(x);
-    stdX = std(x);
-    x_norm = x- meanX;
-    x_out = x_norm(x_norm>3*stdX | x_norm< -3*stdX);
-    idxOutliersCurrent = find(x_norm>3*stdX | x_norm< -3*stdX);
-    idxOutliers = union(idxOutliers, idxOutliersCurrent);
-end
+ypred = sigmoid(tX*beta)>0.5;
 
+tX1 = tX(ypred == 1,:);
 
+tX2 = tX(ypred == 0,:);
 
+scatter3(tX1(:,1),tX1(:,2),tX1(:,3));
+
+hold on
+
+scatter3(tX2(:,1),tX2(:,2),tX2(:,3));
+
+hold off
 

### 2015-10-28 Roman code/dataAnalysis1.m

@@ -1,0 +1,52 @@
+% Load data
+clear all;
+load('Rome_regression.mat');
+
+% plot data histogram
+%figure(1);
+%hist(y_train,100);
+%title('y train')
+
+% plot all X dimensions
+%figure(2);
+%for i = 1:77
+%    subplot(10,8,i);
+%    hist(X_train(:,i),100);
+%end
+    
+
+%figure(3);
+%boxplot(X_train(:,:));
+
+%% Classified data
+% Binary data [0,1]: idx = 9,39,58,61
+idx_zero = [9 39 58 61];
+% Classes [0,2] : idx = 73
+% Classes [1,2] : idx = 3, 47
+% Classes [0, 2, 1] : idx = 2, 57, 67
+% Classes [0, 1, 2, 3] : idx = 7, 12
+idx_classes = [73 3 47 2 57 67 7 12];
+%% 
+% Clean classified data
+idx_X_train = 1:size(X_train,2);
+idx_X_train([idx_zero idx_classes]) = [];
+X_train_cleaned = X_train(:,idx_X_train);
+
+% Find outliers
+idxOutliers = [];
+for i = 1:size(X_train,2)
+    if i ~= 66
+       continue;
+    end
+    x = X_train(:,i);
+    meanX = mean(x);
+    stdX = std(x);
+    x_norm = x- meanX;
+    x_out = x_norm(x_norm>3*stdX | x_norm< -3*stdX);
+    idxOutliersCurrent = find(x_norm>3*stdX | x_norm< -3*stdX);
+    idxOutliers = union(idxOutliers, idxOutliersCurrent);
+end
+
+
+
+

### 2015-10-28 Roman code/dataDescription1.m

@@ -1,0 +1,39 @@
+% Load data
+clear all;
+load('Rome_classification.mat');
+
+%% Plot y histogramm
+figure(1);
+hist(y_train,100);
+
+%% Plot X_train histograms
+figure(2);
+for i = 1:size(X_train,2)
+    subplot(7,5,i);
+    hist(X_train(:,i), 100);
+    title(i);
+end
+
+%% Plot two random variables
+figure(3);
+for i = 1: 12
+    h = randi(34);
+    j = randi(34);
+    subplot(4,3,i);
+    plot(X_train((y_train==1),h),X_train((y_train==1),j), 'k+', 'LineWidth',2, 'MarkerSize',7);
+    hold on;
+    plot(X_train((y_train==-1),h), X_train((y_train==-1),j), 'ko','MarkerFaceColor', 'y', 'MarkerSize',7);
+end
+
+%% Plot boxplot
+figure(4);
+subplot(2,1,1);
+boxplot(X_train);
+title('As-is data');
+subplot(2,1,2);
+X = normalize(X_train);
+boxplot(normalize(X_train));
+title('Normalized data');
+
+
+

### 2015-10-28 Roman code/logLoss.m

@@ -1,6 +1,7 @@
-function logLoss = logLoss(y, y_pred)
+function loss = logLoss(y, y_pred)
 % Log Loss error calculation
-    m = length(y);
-    logLoss = 1/m.*sum(y.*log(y_pred) + (1-y).*log(1-y_pred));
+    m = size(y,1);
+    y_pred = y_pred + 1e-5;    
+    loss = (-1/m) * sum(y.*log(y_pred) + (1.-y).*log(1.-y_pred));
 end
 

### 2015-10-28 Roman code/logisticRegression.m

@@ -1,32 +1,31 @@
 function beta = logisticRegression(y,tX,alpha)
 %Logistic regression using Newton's method.
 
-maxIter = 3; % number of iteration
+maxIter = 10; % number of iteration
 beta = zeros(size(tX,2),1); % initialize beta
-convergence = 1e-5; % convergence threshold
+%beta = randn(size(tX,2),size(y,2));
+convergence = 1.0e-05; % convergence threshold
 m = length(y); % number of data vectors
 
- for i = 1:maxIter
-     
+for i = 1:maxIter
      % Calculate gradient
-     g = (1/m).*tX'*(sigmoid(tX*beta)-y);
-     
+     g = 1./m * tX' *(sigmoid(tX*beta) - y);
+          
      % Calculate Hessian
-     S = diag(sigmoid(tX*beta).*(1-sigmoid(tX*beta)));
-     %H = (1/m).* tX'*S*tX;
-     H = (1/m).*tX'*S*tX;
-     
+     ybar = sigmoid(tX*beta);
+     S = ybar .* (1. - ybar) + 1e-5;
+     H = 1/m * tX' * diag(S) * tX;
+              
      % Check convergence
-     gradient = g'*g
-     if g'*g < convergence; break; end
+     if g'*g < convergence; break; end;
      
      % Calculate cost function for debugging
-     L = (1/m).* sum(-y.*sigmoid(tX*beta)-(1-y).*(1 - sigmoid(tX*beta)));
-     
-     % Calculate beta
-     beta = beta - alpha*H\g;
+     L =  -1/m *( y'*tX*beta - sum(log(1 + exp(tX*beta))) );
           
- end
-
+     % Calculate beta
+     beta = beta - alpha.*H\g;
+end
+L
+i
 end
 

### 2015-10-28 Roman code/main_classification.m

@@ -24,76 +24,75 @@ for i = 1:size(X_train, 2)
     X_train(:,i) = X_train(:,i)./stdX(i);
 end
 
+alpha = 0.0001;
+beta = logisticRegression(y_train,X_train,alpha);
 %%
-% split data in K fold (we will only create indices)
-setSeed(1);
-K = 4;
-N = size(y_train,1);
-idx = randperm(N);
-Nk = floor(N/K);
-for k = 1:K
-	idxCV(k,:) = idx(1+(k-1)*Nk:k*Nk);
-end
+% % split data in K fold (we will only create indices)
+% setSeed(1);
+% K = 4;
+% N = size(y_train,1);
+% idx = randperm(N);
+% Nk = floor(N/K);
+% for k = 1:K
+% 	idxCV(k,:) = idx(1+(k-1)*Nk:k*Nk);
+% end
 
-%%
-% K-fold cross validation
-for k = 1:1
-    % get k'th subgroup in test, others in train
-	idxTe = idxCV(k,:);
-	idxTr = idxCV([1:k-1 k+1:end],:);
-	idxTr = idxTr(:);
-	yTe = y_train(idxTe);
-	XTe = X_train(idxTe,4);
-	yTr = y_train(idxTr);
-	XTr = X_train(idxTr,4);
-        
-	% form tX Warning: Matrix is singular, close to singular or badly scaled. Results may be inaccurate
-    tXTr = [ones(length(yTr), 1) XTr];
-    tXTe = [ones(length(yTe), 1) XTe];
-	
-    % Logistic regression training
-<<<<<<< HEAD
-    beta = penLogisticRegression(yTr, tXTr,0.1,10000);
-=======
-    beta = logisticRegression(yTr, tXTr, 0.01);
->>>>>>> 195ef5f8e92334829e53f06bd62df7d21f05ee78
-    
-    % Make prediction 
-    yTrPred = predClass(tXTr,beta);
-    yTePred = predClass(tXTe,beta);
-    
-    % training and test errors calculation
-    % RMSE error
-    rmseTrSub(k) = rmse(yTr,sigmoid(tXTr*beta));
-	rmseTeSub(k) = rmse(yTe,sigmoid(tXTe*beta)); 
-    % 1/0 loss error
-    binTrSub(k) = binLoss(yTr,yTrPred);
-    binTeSub(k) = binLoss(yTe,yTePred);
[…]

### 2015-10-28 Roman code/main_classification2.m

@@ -1,0 +1,67 @@
+clear all;
+load('Rome_classification');
+
+% Parameters
+alpha = 1.0;
+X = normalize(X_train);
+y_train(y_train == -1) = 0;
+
+%% split data in K fold (we will only create indices)
+setSeed(1);
+K = 4;
+N = size(y_train,1);
+idx = randperm(N);
+Nk = floor(N/K);
+for k = 1:K
+    idxCV(k,:) = idx(1+(k-1)*Nk:k*Nk);
+end
+
+%% Features engineering
+X = normalize(X_train);
+classifiedFeatures = [3 16 25 28 34];
+
+X = poly(X,2);
+X = normalize(X);
+
+%% K-fold cross validation
+tic
+for k = 1:K
+    idxTe = idxCV(k,:);
+    idxTr = idxCV([1:k-1 k+1:end],:);
+	idxTr = idxTr(:);
+    yTe = y_train(idxTe);
+    yTr = y_train(idxTr);
+    XTe = X(idxTe,:);
+    XTr = X(idxTr,:);
+    tXTe = [ones(size(XTe,1),1) XTe];
+    tXTr = [ones(size(XTr,1),1) XTr];
+    
+
+    % Model train
+    beta = penLogisticRegression(yTr, tXTr, alpha, 0.1);
+
+    % Prediction
+    trPr = sigmoid(tXTr*beta);
+    tePr = sigmoid(tXTe*beta);
+    trBin = +(trPr >= 0.5);
+    teBin = +(tePr >= 0.5);
+
+    % Error calculation
+    rmseTr(k) = rmse(yTr, trPr);
+    rmseTe(k) = rmse(yTe, tePr);
+    binLossTr(k) = binLoss(yTr, trBin);
+    binLossTe(k) = binLoss(yTe, teBin);
+    logLossTr(k) = logLoss(yTr, trBin);
+    logLossTe(k) = logLoss(yTe, teBin);
+end
+toc
+%% Mean error calculation
+rmseTr = mean(rmseTr);
+rmseTe = mean(rmseTe);
+binLossTr = mean(binLossTr);
+binLossTe = mean(binLossTe);
+logLossTr = mean(logLossTr);
+logLossTe = mean(logLossTe);
+[rmseTr; rmseTe; binLossTr; binLossTe]
+[logLossTr; logLossTe]
+

### 2015-10-28 Roman code/main_classification3_lambda_choice.m

@@ -1,0 +1,78 @@
+clear all;
+load('Rome_classification');
+
+% Parameters
+alpha = 1.0;
+
+X = normalize(X_train);
+y_train(y_train == -1) = 0; 
+
+%% split data in K fold (we will only create indices)
+setSeed(1);
+K = 4;
+N = size(y_train,1);
+idx = randperm(N);
+Nk = floor(N/K);
+for k = 1:K
+    idxCV(k,:) = idx(1+(k-1)*Nk:k*Nk);
+end
+
+%% Features engineering
+X = normalize(X_train);
+classifiedFeatures = [3 16 25 28 34];
+
+X = poly(X,3);
+X = normalize(X);
+
+%% K-fold cross validation
+tic
+lambdas = logspace(-4,2,100);
+for i = 1:length(lambdas)
+    % K-fold cross validation
+    for k = 1:K
+        idxTe = idxCV(k,:);
+        idxTr = idxCV([1:k-1 k+1:end],:);
+        idxTr = idxTr(:);
+        yTe = y_train(idxTe);
+        yTr = y_train(idxTr);
+        XTe = X(idxTe,:);
+        XTr = X(idxTr,:);
+        tXTe = [ones(size(XTe,1),1) XTe];
+        tXTr = [ones(size(XTr,1),1) XTr];
+    
+        % Model train
+        beta = penLogisticRegression(yTr, tXTr, alpha,lambdas(i));
+
+        % Prediction
+        trPr = sigmoid(tXTr*beta);
+        tePr = sigmoid(tXTe*beta);
+        %trBin(trPr > 0.5) = 1;
+        %teBin(tePr > 0.5) = 1;
+
+        % Error calculation
+        rmseTrK(k) = rmse(yTr, trPr);
+        rmseTeK(k) = rmse(yTe, tePr);
+    end
+
+    % Mean error calculation
+    rmseTr(i) = mean(rmseTrK);
+    rmseTe(i) = mean(rmseTeK);
+end
+toc
+%% Error calculation
+figure(1);
+subplot(2,1,1);
+semilogx(lambdas, rmseTr,'color','b', 'LineWidth',2);
+ylabel('Train RMSE','fontsize',12,'fontname','Helvetica')
+xlabel('lambda','fontsize',12,'fontname','Helvetica')
+grid on;
+subplot(2,1,2);
+semilogx(lambdas, rmseTe,'color','r', 'LineWidth',2);
[…]

### 2015-10-28 Roman code/normalize.m

@@ -1,0 +1,6 @@
+function xo = normalize( xi )
+%NORMALIZE Summary of this function goes here
+%   Detailed explanation goes here
+    xo = (xi-ones(size(xi,1),1)*mean(xi,1))./(ones(size(xi,1),1)*std(xi,0,1));
+end
+

### 2015-10-28 Roman code/penLogisticRegression.m

@@ -5,33 +5,29 @@ function beta = penLogisticRegression(y, tX, alpha, lambda)
 maxIter = 1000; % number of iteration
 beta = zeros(size(tX,2),1); % initialize beta
 convergence = 1e-5; % convergence threshold
-m = length(y); % number of data vectors
-
+m = size(y,1); % number of data vectors
 for i = 1:maxIter
-     % Calculate gradient
-     g = (1/m).*(tX'*(sigmoid(tX*beta)-y));
      
-     % Calculate Hessian (with implemented regularization)
-     S = diag(sigmoid(tX*beta).*(1-sigmoid(tX*beta)));
+     % Calculate gradient
+     g = 1./m * tX' *(sigmoid(tX*beta) - y);
+     
+     % Calculate Hessian
+     ybar = sigmoid(tX*beta);
+     S = ybar .* (1. - ybar) + 1e-5;
      Id = eye(size(tX,2));
      Id(1,1) = 0;
-     %H = (1/m).* tX'*S*tX + (lambda/m).*Id;
-     H = tX'*S*tX + (lambda/m).*Id;
+     H = 1./m * tX' * diag(S) * tX + (lambda/m).*Id;
+    
+     % Check convergence
+     if g'*g < convergence; break; end;
      
      % Calculate cost function for debugging
-     L = (1/m).* sum(-y.*sigmoid(tX*beta)-(1-y).*(1 - sigmoid(tX*beta))); 
+     L =  -1/m *( y'*tX*beta - sum(log(1 + exp(tX*beta))) );
      
      % Calculate beta
-<<<<<<< HEAD
      beta = beta - alpha.*H\g;
-=======
-     beta = beta - alpha*H\g;
->>>>>>> 195ef5f8e92334829e53f06bd62df7d21f05ee78
-     
-     % Check convergence
-     gradinent = g'*g
-     if g'*g < convergence; break; end
 end
-
+L
+i
 end
 

### 2015-10-28 Roman code/poly.m

@@ -1,4 +1,5 @@
 function tXpoly = poly(tX,n)
+    tXpoly = [];
     for d = 1:size(tX,2)
         tXpolyFeature = zeros(size(tX,1),n);
         tXpolyFeature(:,1) = tX(:,d);

### 2015-10-28 Roman code/predClass.m

@@ -1,15 +1,0 @@
-function yPred = predClass(tX, beta)
-% Make prediction for logistic regression
-m = length(tX);
-yPred = zeros(m,1);
-yLog = sigmoid(tX*beta);
-for i = 1:m
-    if yLog(i) > 0.5
-        yPred(i,1) = 1;
-    end
-end
-
-
-
-end
-

### 2015-10-28 Roman code/rmse.m

@@ -1,7 +1,7 @@
-function rmse = computeRMSE(y,y_pred)
+function rmse = rmse(y,y_pred)
 % Compute RMSE for regression problem
     m = length(y);
-    mse = (1/m).* sum((y-y_pred).^2);
+    mse = (1/m) * sum((y-y_pred).^2);
     rmse = sqrt(mse);
 end
 

### 2015-10-28 Roman code/sigmoid.m

@@ -1,5 +1,5 @@
-function h = sigmoid(x)
+function s = sigmoid(x)
 % Sigmoid function
-    h =1.0 ./(1.0 + exp(-x));
+    s = 1./(1 + exp(-x));
 end
 

### 2015-10-28 Roman code/testCode.m

@@ -31,7 +31,7 @@ disp('OK!');
 % testing ridge regression
 disp('ridge regression...');
 lambda = 1e-5;
-tBeta = ridgeRegression(y, tX, lambda);
+tBeta = ridgeRegressionR(y, tX, lambda);
 assert(all(abs(tBeta - beta) < eps));
 disp('OK!');
 
@@ -41,7 +41,7 @@ y = (y>0);
 % testing logistic regression
 disp('logistic regression...');
 alpha = 1e-3;
-tBeta = logisticRegression(y,tX,alpha);
+tBeta = logisticRegressionR(y,tX,alpha);
 tY = 1.0 ./ (1.0 + exp(-tX * tBeta)) > 0.5;
 assert(sum(tY ~= y) / size(y,1) < 0.2);
 disp('OK!');

### 2015-10-28 Roman code/tests.m

@@ -1,0 +1,42 @@
+clear all;
+%load('height_weight_gender');
+load('Rome_classification');
+
+% X = [height weight];
+% X = normalize(X);
+% tX = [ones(size(X,1),1) X];
+% y = +gender; 
+
+X = normalize(X_train);
+tX = [ones(size(X_train,1),1) X_train];
+y = (y_train == 1); 
+
+
+%% Data spliting
+[tXTr, yTr, tXTe, yTe] = split(y, tX, 0.8);
+
+%% My impletation
+tic
+beta = penLogisticRegression(yTr, tXTr, 1.0, 0.01);
+toc
+
+%% Builtin solution
+options = optimset('GradObj', 'on', 'MaxIter', 400);
+initial_theta = zeros(size(tXTr,2),1);
+
+%  Run fminunc to obtain the optimal theta
+[theta, cost] = ...
+	fminunc(@(t)(costFunctionRegNg(t, tXTr, yTr,0.01)), initial_theta, options);
+cost
+%% Prediction
+myTe = +sigmoid(tXTe*beta)>=0.5;
+ngTe = +sigmoid(tXTe*theta)>=0.5;
+
+%% Errors calculation
+myRmse = rmse(yTe, sigmoid(tXTe*beta))
+ngRmse = rmse(yTe, sigmoid(tXTe*theta))
+myBin = binLoss(yTe, myTe)
+ngBin = binLoss(yTe, ngTe)
+myLog = logLoss(yTe, sigmoid(tXTe*beta))
+ngLog = logLoss(yTe, sigmoid(tXTe*theta))
+   

### 2015-10-28 dataAnalysis.m

@@ -5,7 +5,7 @@ tX = normalize(tX);
 
 tY = data.y_train;
 
-alpha = 0.00001;
+alpha = 0.07;
 
 beta = logisticRegression(tY,tX,alpha);
 

### 2015-10-28 logisticRegression.m

@@ -2,7 +2,7 @@ function [ beta ] = logisticRegression( tY,tX,alpha )
 %LOGISTICREGRESSION Summary of this function goes here
 %   Detailed explanation goes here
     cvflag = 1e-5;
-    maxiter = 100000;
+    maxiter = 100;
     beta = randn(size(tX,2),size(tY,2));
     iter = 0;
     g = ones(size(tX,2),size(tY,2));
@@ -10,9 +10,11 @@ function [ beta ] = logisticRegression( tY,tX,alpha )
         iter = iter + 1;
         Cost = compute01Cost(tY,tX,beta);
         g = computeGradLR(tY,tX,beta);
+        norm(g)
         H = computeHessianLR(tY,tX,beta);
-        Cost
+        %Cost
         beta = beta - alpha*H\g;
     end
+    iter
 end
 

### 2015-10-29 getOnehot.m
Reimplemented some missing files. Legend still missing.

@@ -1,0 +1,10 @@
+function out = getOnehot(in,ncat)
+%UNTITLED Summary of this function goes here
+%   Detailed explanation goes here
+    N = size(in,1);
+    out = zeros(N,ncat);
+    for i=1:N
+        out(i,in(i)+1) = 1;
+    end
+end
+

### 2015-10-29 learningCurves.m

@@ -1,0 +1,33 @@
+run('loadAndTreat');
+
+nsample = 30;
+N = size(xTr,1);
+p = 0.8;
+lambda=1e-6;
+trxTr = xTr(1:ceil(p*N),:);
+texTr = xTr(ceil(p*N)+1:N,:);
+tryTr = yTr(1:ceil(p*N),:);
+teyTr = yTr(ceil(p*N)+1:N,:);
+props = linspace(0,1,100);
+
+trerrs = zeros(nsample,100);
+teerrs = zeros(nsample,100);
+for n=1:nsample
+    for j=1:100 
+        trtrxTr = trxTr(1:ceil(props(j)*p*N),:);
+        trtryTr = tryTr(1:ceil(props(j)*p*N),:);
+        beta = ridgeRegression(trtryTr,trtrxTr,lambda);
+        trerrs(n,j) = computeCostLS(trtryTr,trtrxTr,beta);
+        teerrs(n,j) = computeCostLS(teyTr,texTr,beta);
+    end
+    A = randperm(size(trxTr,1));
+    trxTr = trxTr(A,:);
+    tryTr = tryTr(A,:);
+end
+
+boxplot(trerrs,'plotstyle','compact');
+hold on;
+boxplot(teerrs,'plotstyle','compact','colors','r');
+hold off;
+ylim([0.03,0.07]);
+        

### 2015-10-29 loadAndTreat.m

@@ -1,0 +1,33 @@
+data = load('Rome_regression.mat');
+
+option = 'poly'
+
+xTr = data.X_train;
+yTr = data.y_train;
+xTe = data.X_test;
+
+switch option
+    case 'poly'
+        degree = 10;
+        x49 = xTr(:,49);
+        x66 = xTr(:,66);
+        xTr = [];
+        for k=1:degree
+            xTr = [xTr x49.^k];
+            xTr = [xTr x66.^k];
+        end
+    otherwise
+        catf = [2 3 7 12 47 57 61 67 73];
+        catnf = [3 3 4 4 3 4 4 3 3];
+
+        xcat = xTr(:,catf);
+
+        xTr(:,catf) = [];
+
+        for i=1:size(catf,2)
+            xTr = [xTr getOnehot(xcat(:,i),catnf(i))];
+        end
+end
+
+xTr = normalize(xTr);
+yTr = normalize(yTr);

### 2015-10-29 performRegression.m

@@ -1,0 +1,19 @@
+run('loadAndTreat');
+
+nlambda = 500;
+lambda = logspace(-8,3,nlambda);
+option = 'kFold';
+trErs = [];
+teErs = [];
+k = 4;
+
+for j=1:nlambda
+    [teEr,trEr] = kFold(yTr,xTr,k,@(y,x) ridgeRegression(y,x,lambda(j)),@computeCostLS);
+    trErs = [trErs trEr];
+    teErs = [teErs teEr];
+end
+
+subplot(2,1,1);
+semilogx(lambda,trErs);
+subplot(2,1,2);
+semilogx(lambda,teErs,'r');

### 2015-10-31 report/paper.tex
Changes to the figures

@@
 \author{
+Roman Shirochenko\\
 Corentin Tallec\\
-Roman Shirochenko\\
 EPFL \\
+\texttt{roman.shirochenko@epfl.ch}\\
 \texttt{corentin.tallec@epfl.ch} \\
@@
 $10^3$, with 500 points. The results of the ridge regression are already
-satisfactory, with an error of approximately $0.105\pm0.001$ on the training set
-and $0.115\pm0.001$ on the test set, that is approximately one tenth of the
+satisfactory, with an error of approximately $0.32\pm0.005$ on the training set
+and $0.34\pm0.005$ on the test set, that is approximately one third of the
 variance of the output variable.

### 2015-10-31 computeCostLS.m

@@ -3,6 +3,6 @@ function cost = computeCostLS( y,tX,beta )
 %   Detailed explanation goes here
     e = y - tX*beta;
     N = size(y,1);
-    cost = e'*e/(2*N);
+    cost = sqrt(e'*e/(2*N));
 end
 

### 2015-10-31 dataAnalysisRegression.m

@@ -2,7 +2,7 @@ data = load('Rome_regression');
 trX = data.X_train;
 trY = data.y_train;
 teX = data.X_test;
-label='boxplot'
+label='cor1'
 N = size(trY,1);
 f = size(trX,2);
 
@@ -11,12 +11,13 @@ switch label
         boxplot(trX);
         title('Box plot for train data');
         set(gca,'xtick',[]);
-        xlabel('Features');
+        hx = xlabel('Features');
+        hy = ylabel('');
     case 'yhist'
         hist(trY,sqrt(N));
         title('Y training set histogram');
-        xlabel('Y value');
-        ylabel('Frequency');
+        hx = xlabel('Y value');
+        hy = ylabel('Frequency');
     case 'xhist'
         for i=1:f
             subplot(ceil(sqrt(f)),ceil(sqrt(f)),i);
@@ -28,12 +29,22 @@ switch label
     case 'xhist49'
         hist(trX(:,49),sqrt(N));
         title('Feature 49');
-        xlabel('Feature value');
-        ylabel('Frequency');
+        hx = xlabel('Feature value');
+        hy = ylabel('Frequency');
     case 'xhist66'
         hist(trX(:,66),sqrt(N));
         title('Feature 66');
-        xlabel('Feature value');
-        ylabel('Frequency');
+        hx = xlabel('Feature value');
+        hy = ylabel('Frequency');
+    case 'cor1'
+        scatter(trX(:,1),trY);
+        title('Y vs first feature');
+        hx = xlabel('First feature');
+        hy = ylabel('Y');
 end
+
+% the following code makes the plot look nice and increase font size etc.
+set(gca,'fontsize',20,'fontname','Helvetica','box','off','tickdir','out','ticklength',[.02 .02],'xcolor',0.5*[1 1 1],'ycolor',0.5*[1 1 1]);
+set([hx; hy],'fontsize',18,'fontname','avantgarde','color',[.3 .3 .3]);
+grid on;
 

### 2015-10-31 learningCurves.m

@@ -25,9 +25,23 @@ for n=1:nsample
     tryTr = tryTr(A,:);
 end
 
-boxplot(trerrs,'plotstyle','compact');
+figure(1);
+boxplot(trerrs,'plotstyle','compact',...
+    'symbol','k.',...
+    'medianstyle','line');
 hold on;
-boxplot(teerrs,'plotstyle','compact','colors','r');
+boxplot(teerrs,'plotstyle','compact',...
+    'symbol','k.',...
+    'medianstyle','line','colors','r');
 hold off;
-ylim([0.03,0.07]);
+
+ylim([0.15 0.3]);
+
+hx = xlabel('Train data proportion');
+hy = ylabel('RMSE');
+
+% the following code makes the plot look nice and increase font size etc.
+set(gca,'fontsize',20,'fontname','Helvetica','box','off','tickdir','out','ticklength',[.02 .02],'xcolor',0.5*[1 1 1],'ycolor',0.5*[1 1 1]);
+set([hx; hy],'fontsize',18,'fontname','avantgarde','color',[.3 .3 .3]);
+grid on;
         

### 2015-10-31 loadAndTreat.m

@@ -1,6 +1,6 @@
 data = load('Rome_regression.mat');
 
-option = 'poly'
+option = 'poly';
 
 xTr = data.X_train;
 yTr = data.y_train;
@@ -11,23 +11,36 @@ switch option
         degree = 10;
         x49 = xTr(:,49);
         x66 = xTr(:,66);
+        tex49 = xTe(:,49);
+        tex66 = xTe(:,66);
         xTr = [];
+        xTe = [];
         for k=1:degree
             xTr = [xTr x49.^k];
             xTr = [xTr x66.^k];
+            xTe = [xTe tex49.^k];
+            xTe = [xTe tex66.^k];
         end
     otherwise
         catf = [2 3 7 12 47 57 61 67 73];
         catnf = [3 3 4 4 3 4 4 3 3];
 
         xcat = xTr(:,catf);
+        texcat = xTe(:,catf);
 
         xTr(:,catf) = [];
+        xTe(:,catf) = [];
 
         for i=1:size(catf,2)
             xTr = [xTr getOnehot(xcat(:,i),catnf(i))];
+            xTe = [xTe getOnehot(texcat(:,i),catnf(i))];
         end
 end
-
+nf = size(xTr,2);
+xstd = std(xTr,1);
+xmean = mean(xTr,1);
+ystd = std(yTr,1);
+ymean = mean(yTr,1);
 xTr = normalize(xTr);
 yTr = normalize(yTr);
+xTe = normalize(xTe);

### 2015-10-31 performRegression.m

@@ -15,5 +15,19 @@ end
 
 subplot(2,1,1);
 semilogx(lambda,trErs);
+hx = xlabel('lambda');
+hy = ylabel('RMSE');
+title('Training error');
+ylim([0.3 0.4]);
+set(gca,'fontsize',20,'fontname','Helvetica','box','off','tickdir','out','ticklength',[.02 .02],'xcolor',0.5*[1 1 1],'ycolor',0.5*[1 1 1]);
+set([hx; hy],'fontsize',18,'fontname','avantgarde','color',[.3 .3 .3]);
+grid on;
 subplot(2,1,2);
 semilogx(lambda,teErs,'r');
+hx = xlabel('lambda');
+hy = ylabel('RMSE');
+title('Test error');
+ylim([0.3 0.4]);
+set(gca,'fontsize',20,'fontname','Helvetica','box','off','tickdir','out','ticklength',[.02 .02],'xcolor',0.5*[1 1 1],'ycolor',0.5*[1 1 1]);
+set([hx; hy],'fontsize',18,'fontname','avantgarde','color',[.3 .3 .3]);
+grid on;

### 2015-10-31 prediction.m

@@ -1,0 +1,16 @@
+run('loadAndTreat');
+nsample = 30;
+lambda = 1e-6;
+k = 4;
+teErrs = zeros(nsample,1);
+for i=1:nsample
+    A = randperm(size(xTr,1));
+    xTr = xTr(A,:);
+    yTr = yTr(A,:);
+    [teErr,trErr] = kFold(yTr,xTr,k,@(y,x) ridgeRegression(y,x,lambda),@computeCostLS);
+    teErrs(i,1) = teErr;
+end
+hist(teErrs);
+
+beta = ridgeRegression(yTr,xTr,lambda);
+pred = xTe*beta*ystd + ymean

### 2015-10-31 prediction.m
Added the results files

@@ -8,9 +8,15 @@ for i=1:nsample
     xTr = xTr(A,:);
     yTr = yTr(A,:);
     [teErr,trErr] = kFold(yTr,xTr,k,@(y,x) ridgeRegression(y,x,lambda),@computeCostLS);
-    teErrs(i,1) = teErr;
+    teErrs(i,1) = teErr*ystd;
 end
-hist(teErrs);
+mean(teErrs)
+std(teErrs)
 
 beta = ridgeRegression(yTr,xTr,lambda);
-pred = xTe*beta*ystd + ymean
+pred = xTe*beta*ystd + ymean;
+teerror = ['rmse', num2str(mean(teErrs))];
+
+f = fopen('test_errors_regression.csv','wt')
+fprintf(f,strcat('rmse,',num2str(mean(teErrs))));
+csvwrite('predictions_regression.csv',pred);

### 2015-10-31 report/paper.tex
Last changes to the report

@@
 and $0.34\pm0.005$ on the test set, that is approximately one third of the
-variance of the output variable.
+standard deviation of the output variable.
 
@@
 As expected, both training error and test error decrease as the proportion of
-training data increases. The test error obtained seems to exhibit a quit
+training data increases. The test error obtained seems to exhibit a quite
 satisfactory behaviour.
@@
 
+\subsection{Regression Summary}
+While simple ridge regression with an appropriately optimized lambda already
+gave decent results on our dataset, we obtained better results using polynomial
+transformations of well chosen features. Doing so, we obtain an error that is
+approximately one fifth of the standard deviation.
+
 \end{document}

### 2015-10-31 prediction.m

@@ -11,8 +11,7 @@ for i=1:nsample
     teErrs(i,1) = teErr*ystd;
 end
 mean(teErrs)
-std(teErrs)
-
+ystd
 beta = ridgeRegression(yTr,xTr,lambda);
 pred = xTe*beta*ystd + ymean;
 teerror = ['rmse', num2str(mean(teErrs))];

### 2015-10-31 report/paper.tex
Added all files and final folder for upload with results

--- 2015-10-27 report/paper.tex
@@
-\documentclass{article} % For LaTeX2e
+\documentclass{article} % For LaTeX2
+
 % We will use NIPS submission format
@@
 
-\title{Project-I by Group CITYNAME}
+\title{Project-I by Group ROME}
 
 \author{
+Roman Shirochenko\\
 Corentin Tallec\\
-Roman Shirochenko\\
 EPFL \\
+\texttt{roman.shirochenko@epfl.ch}\\
 \texttt{corentin.tallec@epfl.ch} \\
@@
 
-The results described here are summerized in the Figure \ref{fig:ridgeCurve}.
+The results described here are summarized in the Figure \ref{fig:ridgeCurve}.
 
@@
 As notified in a previous section, among all the feature at our disposal,
-features 49 and 66 show distributions that ressemble the ditribution of the
+features 49 and 66 show distributions that resemble the distribution of the
 output variable.
 
-This led us to try and regress only on thos two features and polynomial
+This led us to try and regress only on those two features and polynomial
 transformations of them. Using only those two features, we get an RMSE error
@@
 $k=4$. The results are summarized in Figure \ref{fig:polyridgeCurve}, where, as
-previously mentionned, we used as features only the two features and their
+previously mentioned, we used as features only the two features and their
 polynomial transformation with degree ranging from 1 to 10.
@@
 
+\section{Logistic Regression}
+\subsection{Data Description}
+Our second data set consisted of a train set, with an input and an output binary
+variable respectively $\mathbf{X}$ and $\mathbf{y}$, composed of $N=1500$ data
+samples, and a test set with only $\mathbf{X}$ variable observed and $N=1500$
+data samples. $\mathbf{X}$ spans over $D=34$ features, among which 5 of them are
+categorical, with number of categories ranging from 1 to 5.
+
+Our goal is to produce binary classification predictions for test examples and approximate the
+test error.
+
+\subsection{Data visualization and cleaning}
+Firstly, the data was visualized. The first thing we
+did was to obtain a visualization of the data, showed in \ref{fig:boxplotX} and
+\ref{fig:histY}. Those two figures led us to normalize both the input and the
+output variable (normalizing the output variable allow us to have results that are
+easier to interpret, as they are to be compared to values close to 1). No
+obvious outliers are noticeable on the $\mathbf{y}$ histogram.
+
+We performed featurewise analysis to notice that most of our real valued
+variables have nearly gaussian distributions (or at least unimodal
+distributions), which is not the case of our output variable. Figure
+\ref{fig:cor1} shows how the output variable evolves as a function of the first
+feature. This behaviour is what we notice on most variables, and thus shows no
+clear correlations.
 \end{document}

### 2015-10-31 Roman code/computeCostLR.m

@@ -1,0 +1,7 @@
+function cost = computeCostLR(y,tX,beta)
+%COMPUTECOSTLR Summary of this function goes here
+%   Detailed explanation goes here
+    N = size(y,1);
+    cost =  -1/N *( y'*tX*beta - sum(log(1 + exp(tX*beta))) );
+end
+

### 2015-10-31 Roman code/computeGradLR.m

@@ -1,0 +1,7 @@
+function [ g ] = computeGradLR( tY, tX, beta )
+%COMPUTEGRADLR Summary of this function goes here
+%   Detailed explanation goes here
+    N = size(tX,1);
+    g = 1./N * tX' * (sigmoid(tX*beta) - tY);
+end
+

### 2015-10-31 Roman code/dataDescription1.m

@@ -4,8 +4,14 @@ load('Rome_classification.mat');
 
 %% Plot y histogramm
 figure(1);
-hist(y_train,100);
-
+xbins2 = [-2 -1 0 1 2];
+hist(y_train, xbins2);
+hy = ylabel('Frequency');
+hx = xlabel('y');
+set(gca,'fontsize',20,'fontname','Helvetica','box','off','tickdir','out','ticklength',[.02 .02],'xcolor',0.5*[1 1 1],'ycolor',0.5*[1 1 1]);
+set([hx; hy],'fontsize',18,'fontname','avantgarde','color',[.3 .3 .3]);
+grid on;
+print('HistX','-dpng')
 %% Plot X_train histograms
 figure(2);
 for i = 1:size(X_train,2)
@@ -27,13 +33,13 @@ end
 
 %% Plot boxplot
 figure(4);
-subplot(2,1,1);
 boxplot(X_train);
-title('As-is data');
-subplot(2,1,2);
-X = normalize(X_train);
-boxplot(normalize(X_train));
-title('Normalized data');
+hy = ylabel('');
+hx = xlabel('X input variables');
+set(gca,'fontsize',12,'fontname','Helvetica','box','off','tickdir','out','ticklength',[.02 .02],'xcolor',0.5*[1 1 1],'ycolor',0.5*[1 1 1]);
+set([hx; hy],'fontsize',12,'fontname','avantgarde','color',[.3 .3 .3]);
+grid on;
+print('boxplotX','-dpng')
 
 
 

### 2015-10-31 Roman code/learningCurves.m

@@ -1,0 +1,51 @@
+clear all;
+load('Rome_classification');
+
+classifiedFeatures = [3 16 25 28 34];
+classifiedRows = [];
+for i = 1:length(classifiedFeatures)
+    classifiedRows = [classifiedRows dummyvar(X_train(:,classifiedFeatures(i)))];
+end
+X_train(:,classifiedFeatures) = [];
+X_train = [X_train classifiedRows];
+X = poly(X_train,3);
+X = normalize(X);
+y_train(y_train == -1) = 0;
+
+xTr = X;
+yTr = y_train;
+
+%% Plot learning curve
+tic
+nsample = 30;
+N = size(xTr,1);
+p = 0.8;
+lambda=0.3;
+alpha = 1.0;
+trxTr = xTr(1:ceil(p*N),:);
+texTr = xTr(ceil(p*N)+1:N,:);
+tryTr = yTr(1:ceil(p*N),:);
+teyTr = yTr(ceil(p*N)+1:N,:);
+props = linspace(0,1,100);
+
+trerrs = zeros(nsample,100);
+teerrs = zeros(nsample,100);
+for n=1:nsample
+    for j=1:100 
+        trtrxTr = trxTr(1:ceil(props(j)*p*N),:);
+        trtryTr = tryTr(1:ceil(props(j)*p*N),:);
+        beta = penLogisticRegression(trtryTr,trtrxTr,alpha, lambda);
+        trerrs(n,j) = computeCostLR(trtryTr,trtrxTr,beta);
+        teerrs(n,j) = computeCostLR(teyTr,texTr,beta);
+    end
+    A = randperm(size(trxTr,1));
+    trxTr = trxTr(A,:);
+    tryTr = tryTr(A,:);
+end
+toc
+%%
+figure(1);
+boxplot(trerrs,'plotstyle','compact');
+figure(2)
+boxplot(teerrs,'plotstyle','compact','colors','r');
+hold off;

### 2015-10-31 Roman code/logLoss.m

@@ -1,7 +1,7 @@
-function loss = logLoss(y, y_pred)
+function loss = logLoss(y_t, y_pred)
 % Log Loss error calculation
-    m = size(y,1);
-    y_pred = y_pred + 1e-5;    
-    loss = (-1/m) * sum(y.*log(y_pred) + (1.-y).*log(1.-y_pred));
+    y_pred(y_pred > 0.99999) = 0.99999;  
+    m = size(y_t,1);
+    loss = (-1/m) * sum(y_t.*log(y_pred) + (1-y_t).*log(1-y_pred));
 end
 

### 2015-10-31 Roman code/main_classification2.m

@@ -1,9 +1,16 @@
 clear all;
 load('Rome_classification');
+%% Dummy encoding
+classifiedFeatures = [3 16 25 28 34];
+classifiedRows = [];
+for i = 1:length(classifiedFeatures)
+    classifiedRows = [classifiedRows dummyvar(X_train(:,classifiedFeatures(i)))];
+end
+X_train(:,classifiedFeatures) = [];
+X_train = [X_train classifiedRows];
 
-% Parameters
-alpha = 1.0;
-X = normalize(X_train);
+X = poly(X_train,3);
+X = normalize(X);
 y_train(y_train == -1) = 0;
 
 %% split data in K fold (we will only create indices)
@@ -16,15 +23,12 @@ for k = 1:K
     idxCV(k,:) = idx(1+(k-1)*Nk:k*Nk);
 end
 
-%% Features engineering
-X = normalize(X_train);
-classifiedFeatures = [3 16 25 28 34];
-
-X = poly(X,2);
-X = normalize(X);
-
 %% K-fold cross validation
+alpha = 1.0;
+binLossTe = zeros(K,1);
+binLossTr = zeros(K,1);
 tic
+
 for k = 1:K
     idxTe = idxCV(k,:);
     idxTr = idxCV([1:k-1 k+1:end],:);
@@ -36,9 +40,8 @@ for k = 1:K
     tXTe = [ones(size(XTe,1),1) XTe];
     tXTr = [ones(size(XTr,1),1) XTr];
     
-
     % Model train
-    beta = penLogisticRegression(yTr, tXTr, alpha, 0.1);
+    beta = penLogisticRegression(yTr, tXTr, alpha, 0.3);
 
     % Prediction
     trPr = sigmoid(tXTr*beta);
@@ -47,21 +50,30 @@ for k = 1:K
     teBin = +(tePr >= 0.5);
 
     % Error calculation
-    rmseTr(k) = rmse(yTr, trPr);
-    rmseTe(k) = rmse(yTe, tePr);
     binLossTr(k) = binLoss(yTr, trBin);
     binLossTe(k) = binLoss(yTe, teBin);
-    logLossTr(k) = logLoss(yTr, trBin);
-    logLossTe(k) = logLoss(yTe, teBin);
 end
 toc
-%% Mean error calculation
-rmseTr = mean(rmseTr);
-rmseTe = mean(rmseTe);
-binLossTr = mean(binLossTr);
-binLossTe = mean(binLossTe);
[…]

### 2015-10-31 Roman code/main_classification3_lambda_choice.m

@@ -3,8 +3,18 @@ load('Rome_classification');
 
 % Parameters
 alpha = 1.0;
+%% Dummy encoding
+classifiedFeatures = [3 16 25 28 34];
+classifiedRows = [];
+for i = 1:length(classifiedFeatures)
+    classifiedRows = [classifiedRows dummyvar(X_train(:,classifiedFeatures(i)))];
+end
+X_train(:,classifiedFeatures) = [];
+X_train = [X_train classifiedRows];
 
-X = normalize(X_train);
+X = poly(X_train,3);
+X = normalize(X);
+
 y_train(y_train == -1) = 0; 
 
 %% split data in K fold (we will only create indices)
@@ -17,16 +27,9 @@ for k = 1:K
     idxCV(k,:) = idx(1+(k-1)*Nk:k*Nk);
 end
 
-%% Features engineering
-X = normalize(X_train);
-classifiedFeatures = [3 16 25 28 34];
-
-X = poly(X,3);
-X = normalize(X);
-
 %% K-fold cross validation
 tic
-lambdas = logspace(-4,2,100);
+lambdas = logspace(-3,4,100);
 for i = 1:length(lambdas)
     % K-fold cross validation
     for k = 1:K
@@ -46,33 +49,52 @@ for i = 1:length(lambdas)
         % Prediction
         trPr = sigmoid(tXTr*beta);
         tePr = sigmoid(tXTe*beta);
-        %trBin(trPr > 0.5) = 1;
-        %teBin(tePr > 0.5) = 1;
+        trBin = +(trPr >= 0.5);
+        teBin = +(tePr >= 0.5);
 
         % Error calculation
         rmseTrK(k) = rmse(yTr, trPr);
         rmseTeK(k) = rmse(yTe, tePr);
+        binLossTrK(k) = binLoss(yTr, trBin);
+        binLossTeK(k) = binLoss(yTe, teBin);
     end
 
     % Mean error calculation
     rmseTr(i) = mean(rmseTrK);
     rmseTe(i) = mean(rmseTeK);
+    binLossTr(i) = mean(binLossTrK);
+    binLossTe(i) = mean(binLossTeK);
 end
 toc
 %% Error calculation
+% figure(1);
+% subplot(2,1,1);
+% semilogx(lambdas, rmseTr,'color','b', 'LineWidth',2);
[…]

### 2015-10-31 Roman code/tests.m

@@ -1,5 +1,4 @@
 clear all;
-%load('height_weight_gender');
 load('Rome_classification');
 
 % X = [height weight];
@@ -8,17 +7,34 @@ load('Rome_classification');
 % y = +gender; 
 
 X = normalize(X_train);
+X = poly(X,5);
+X = normalize(X);
 tX = [ones(size(X_train,1),1) X_train];
 y = (y_train == 1); 
 
 
 %% Data spliting
-[tXTr, yTr, tXTe, yTe] = split(y, tX, 0.8);
+
+[tXTr, yTr, tXTe, yTe] = split(y, tX, 0.75);
+
 
 %% My impletation
+
+% tic
+% beta = penLogisticRegression(yTr, tXTr, 1.0, lambda);
+% toc
+lambda = 0.001;
 tic
-beta = penLogisticRegression(yTr, tXTr, 1.0, 0.01);
+beta = penLogisticRegression(yTr, tXTr, 1,lambda);
 toc
+
+%% Prediction
+myTr = +sigmoid(tXTr*beta)>=0.5;
+myTe = +sigmoid(tXTe*beta)>=0.5;
+
+% Errors calculation
+myBinTr = binLoss(yTr, myTr)
+myBinTe = binLoss(yTe, myTe)
 
 %% Builtin solution
 options = optimset('GradObj', 'on', 'MaxIter', 400);
@@ -26,17 +42,5 @@ initial_theta = zeros(size(tXTr,2),1);
 
 %  Run fminunc to obtain the optimal theta
 [theta, cost] = ...
-	fminunc(@(t)(costFunctionRegNg(t, tXTr, yTr,0.01)), initial_theta, options);
+	fminunc(@(t)(costFunctionRegNg(t, tXTr, yTr, lambda)), initial_theta, options);
 cost
-%% Prediction
-myTe = +sigmoid(tXTe*beta)>=0.5;
-ngTe = +sigmoid(tXTe*theta)>=0.5;
-
-%% Errors calculation
-myRmse = rmse(yTe, sigmoid(tXTe*beta))
-ngRmse = rmse(yTe, sigmoid(tXTe*theta))
-myBin = binLoss(yTe, myTe)
-ngBin = binLoss(yTe, ngTe)
-myLog = logLoss(yTe, sigmoid(tXTe*beta))
-ngLog = logLoss(yTe, sigmoid(tXTe*theta))
-   

### 2015-10-31 submit_files_folder/leastSquaresGD.m

@@ -1,0 +1,27 @@
+function beta = leastSquaresGD(y,tX,alpha)
+% Least squares using gradient descent
+% alpha is step size
+
+maxIters = 1000; % number of iterations
+converged = 1e-5; %number for convergence check
+N = length(y); % compute number of data points
+beta = zeros(size(tX,2),1); % initiate beta
+    
+for k = 1:maxIters
+    % compute the error of the model
+    e = y - tX*beta; 
+
+    % compute gradient
+    g = ((-1)/N)*tX'*e;
+
+    % compute cost function
+    L = e'*e/(2*N);
+
+    % update beta
+    beta = beta - alpha .*g;
+
+    % check for convergence
+    if g'*g < converged; break; end;
+end
+end
+

### 2015-10-31 submit_files_folder/logisticRegression.m

@@ -1,9 +1,8 @@
 function beta = logisticRegression(y,tX,alpha)
 %Logistic regression using Newton's method.
 
-maxIter = 10; % number of iteration
+maxIter = 1000; % number of iteration
 beta = zeros(size(tX,2),1); % initialize beta
-%beta = randn(size(tX,2),size(y,2));
 convergence = 1.0e-05; % convergence threshold
 m = length(y); % number of data vectors
 
@@ -25,7 +24,5 @@ for i = 1:maxIter
      % Calculate beta
      beta = beta - alpha.*g;
 end
-L
-i
 end
 

### 2015-10-31 submit_files_folder/penLogisticRegression.m

@@ -2,12 +2,11 @@ function beta = penLogisticRegression(y, tX, alpha, lambda)
 % Penalized logistic regression using Newton's method
 % lambda is the regularization parameter
 
-maxIter = 200; % number of iteration
+maxIter = 1000; % number of iteration
 beta = zeros(size(tX,2),1); % initialize beta
 convergence = 1e-5; % convergence threshold
 m = size(y,1); % number of data vectors
 for i = 1:maxIter
-     
      % Calculate gradient
      g = 1./m * tX' *(sigmoid(tX*beta) - y);
      
@@ -21,12 +20,11 @@ for i = 1:maxIter
      % Check convergence
      if g'*g < convergence; break; end;
      
-     % Calculate cost function for debugging
+     % Calculate cost function 
      %L =  -1/m *( y'*tX*beta - sum(log(1 + exp(tX*beta))) );
      
      % Calculate beta
      beta = beta - alpha.*H\g;
 end
-
 end
 

### 2015-10-31 submit_files_folder/ridgeRegression.m

@@ -1,0 +1,7 @@
+function beta = ridgeRegression(y,tX,lambda)
+% Ridge regression using normal equation
+Id = eye(size(tX,2));
+Id(1) = 0; % To do not regularize beta0
+beta = (tX'*tX + lambda * Id)\tX'*y;
+end
+

### 2015-10-31 submit_files_folder/sigmoid.m

@@ -1,0 +1,5 @@
+function s = sigmoid(x)
+% Sigmoid function
+    s = 1./(1 + exp(-x));
+end
+

### 2015-10-31 submit_files_folder/testCode.m

@@ -57,6 +57,6 @@ disp('OK!');
 
 % code for writing csv files
 tY = 1.0 ./ (1.0 + exp(-tX * tBeta));
-csvwrite('predictions_classification.csv', tY);
+%csvwrite('predictions_classification.csv', tY);
 
 disp('test finished successfully');

### 2015-10-31 report/paper.tex

@@
+<<<<<<< HEAD
 \documentclass{article} % For LaTeX2
@@
 \end{document}
+=======
+\documentclass{article} % For LaTeX2e
+% We will use NIPS submission format
+\usepackage{nips13submit_e,times}
+% for hyperlinks
+\usepackage{hyperref}
+\usepackage{url}
+% For figures
+\usepackage{graphicx} 
+\usepackage{subfigure} 
+% math packages
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amsopn}
+\usepackage{ifthen}
+\usepackage{natbib}
+
+\title{Project-I by Group CITYNAME}
+
+\author{
+Roman Shirochenko\\
+Corentin Tallec\\
+EPFL \\
+\texttt{roman.shirochenko@epfl.ch}\\
+\texttt{corentin.tallec@epfl.ch} \\
+}
+
+% The \author macro works with any number of authors. There are two commands
+% used to separate the names and addresses of multiple authors: \And and \AND.
+%
+% Using \And between authors leaves it to \LaTeX{} to determine where to break
+% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
+% puts 3 of 4 authors names on the first line, and the last on the second
+% line, try using \AND instead of \And before the third author name.
+
+\nipsfinalcopy
+
+\begin{document}
+
+\maketitle
+
+\begin{abstract}
+	In this report we summarize the results we obtained for the first machine
+	learning of the EPFL machine learning course. Facing two datasets, we fit
+	linear models and perform linear regression on the first one, and perform data
+	classification using logistic regression on the second one. For the first
+	dataset, we find that simple linear regression and ridge regression gives
+	decent results, but that we can get better results by isolating one simple
+	feature and performing transformations on that feature.
+\end{abstract}
+
+\section{Linear Regression}
+\subsection{Data Description}
+Our first data set consisted of a train set, with an input and an output
+variable respectively $\mathbf{X}$ and $\mathbf{y}$, composed of $N=2800$ data
+samples, and a test set with only $\mathbf{X}$ variable observed and $N=1200$
+data samples. $\mathbf{X}$ spans over $D=77$ features, among which 9 of them are
+categorical, with number of categories ranging from 2 to 4, 2 of them are binary
+and the other are real-valued.
+
+Our goal is to produce linear predictions for test examples and approximate the
+test error.
+
+\subsection{Data visualization and cleaning}
+The first tasks we performed on the data was to visualize it. The first thing we
+did was to obtain a visualization of the data, showed in \ref{fig:boxplotX} and
+\ref{fig:histY}. Those two figures led us to normalize both the input and the
+output variable (normalizing the output variable allow us to have results that are
+easier to interpret, as they are to be compared to values close to 1). No
+obvious outliers are noticeable on the $\mathbf{y}$ histogram.
+
+We performed featurewise analysis to notice that most of our real valued
+variables have nearly gaussian distributions (or at least unimodal
+distributions), which is not the case of our output variable. Figure
+\ref{fig:cor1} shows how the output variable evolves as a function of the first
+feature. This behaviour is what we notice on most variables, and thus shows no
+clear correlations.
+
+Although for variables 49 and 66 we notice a multimodal distribution that is
+closer to the behaviour of our output variable, as seen in Figure
+\ref{fig:xhist49} and Figure \ref{fig:xhist66}. We thus chose to try and predict
+$\mathbf{y}$ using only those two features, transformed through polynomial
+transformation.
+
+The categorical variables were encoded using dummy encoding, what led to a total
+feature number of 100. As the rank of our dataset is below 100 (it is 70),
+simple linear regression is likely to fail, and we use ridge regression to cope
+for the rank deficiency.
+
+\begin{figure}[!t]
+\center
+\subfigure[Boxplot of $\mathbf{X}$.]{\includegraphics[width=2.5in]{figures/boxplot.jpg} \label{fig:boxplotX}}
+\subfigure[$\mathbf{y}$ as a function of the first feature of $\mathbf{X}$.
+The correlation between the two variables is not
+clear.]{\includegraphics[width=2.5in]{figures/cor1.jpg}\label{fig:cor1}}
+\hfill
+\subfigure[Histogram of
+$\mathbf{y}$. The output distribution is bimodal.]{\includegraphics[width=1.5in]{figures/yhist.jpg} \label{fig:histY}}
+\hspace*{0.3in}
+\subfigure[Histogram of feature 49 of $\mathbf{X}$. We see that the distribution
+	is bimodal, as the output.]{\includegraphics[width=1.5in]{figures/xhist49.jpg}
+\label{fig:xhist49}}
+\hspace*{0.3in}
+\subfigure[Histogram of feature 66 of $\mathbf{X}$. Again, the distribution is
+	bimodal.]{\includegraphics[width=1.5in]{figures/xhist66.jpg}
+\label{fig:xhist66}}
+\caption{}
+\end{figure}
+
+\subsection{Ridge regression}
+As the dimension of the data is inferior to the number of feature, we expect
+least-squares to work quite poorly, and thus decided to apply mostly ridge
+regression.
+
+For our first attempt at predicting the data, we simply used the initial
+features and used kFold with $k=4$ to display the evolution of the RMSE as a
+function of the regularizing lambda both for the training and test data. 
+The lambda we used ranged from $10^{-4}$ to
+$10^3$, with 500 points. The results of the ridge regression are already
+satisfactory, with an error of approximately $0.32\pm0.005$ on the training set
+and $0.34\pm0.005$ on the test set, that is approximately one third of the
+standard deviation of the output variable.
+
+We see that for sufficiently low lambda, both the training and test errors do
+not vary much (there is indeed only a very slight variation of test RMSE, with a
+nearly unnoticeable minimum between $\lambda=10$ and $\lambda=100$).
+
+The results described here are summerized in the Figure \ref{fig:ridgeCurve}.
+
+\subsection{Feature transformations}
+As notified in a previous section, among all the feature at our disposal,
+features 49 and 66 show distributions that ressemble the ditribution of the
+output variable.
+
+This led us to try and regress only on thos two features and polynomial
+transformations of them. Using only those two features, we get an RMSE error
+that is quite close to the one we notified when all the features are used. When
+using polynomial features (up to degree 10), we basically get half the error
+that we got using all the features. Those results were obtained using ridge
+regression. As in the previous section, we obtained the RMSE as a function of
+lambda, calculating both training and test errors using a kFold procedure with
+$k=4$. The results are summarized in Figure \ref{fig:polyridgeCurve}, where, as
+previously mentionned, we used as features only the two features and their
+polynomial transformation with degree ranging from 1 to 10.
+
+This time, the test error varies much more as a function of lambda, and, given
+our curve, the best lambda possible seems to be around $10^{-6}$.
+
+Finally to evaluate our performances more accurately, we plotted learning curves
+for our model, with 20 features obtained by polynomial transformations and a
+lambda of $10^{-6}$. The learning curve is obtained by using $20\%$ of the
+training data to perform tests, while learning the model on a progressively
+increasing portion of the $80\%$ remaining. We monitor the test and training
+error as functions of the proportion of data used to train. The results are
+summarized in Figure \ref{fig:learningCurve}.
+
+As expected, both training error and test error decrease as the proportion of
+training data increases. The test error obtained seems to exhibit a quite
+satisfactory behaviour.
+
+\begin{figure}[!h]
+\center
+\subfigure[Ridge regression on kFold with
+$k=4$.]{\includegraphics[width=2.5in]{figures/traintestRMSE.jpg} \label{fig:ridgeCurve}}
+\hfill
+\subfigure[Ridge regression on kFold with feature transformation and
+	$k=4$.]{\includegraphics[width=2.5in]{figures/traintestRMSEpoly.jpg}
+\label{fig:polyridgeCurve}}
+\subfigure[Learning curve. Blue is training data and red is test
+data.]{\includegraphics[width=4in]{figures/learningcurves.jpg}
+\label{fig:learningCurve}}
+\caption{}
+\end{figure}
+
+\subsection{Regression Summary}
+While simple ridge regression with an appropriately optimized lambda already
+gave decent results on our dataset, we obtained better results using polynomial
+transformations of well chosen features. Doing so, we obtain an error that is
+approximately one fifth of the standard deviation.
+
+\end{document}
+>>>>>>> cad465af211561b5f7c07cebe323918cf8ff2cfd

### 2015-10-31 report/paper.tex
Added Logistic regression section"

--- 2015-10-31 report/paper.tex
@@
 
+\section{Logistic Regression}
+\subsection{Data Description}
+Our second data set consisted of a train set, with an input and an output binary
+variable respectively $\mathbf{X}$ and $\mathbf{y}$, composed of $N=1500$ data
+samples, and a test set with only $\mathbf{X}$ variable observed and $N=1500$
+data samples. $\mathbf{X}$ spans over $D=34$ features, among which 5 of them are
+categorical, with number of categories ranging from 1 to 5.
+
+The categorical variables were encoded using dummy encoding, what led to a total
+feature number of 47. As the rank of our dataset is below 47 (it is 42), we use penalized logistic regression to cope for the rank deficiency.
+
+Our goal is to produce binary classification predictions for test examples and approximate the test error.
+
+\subsection{Data visualization and cleaning}
+Firstly, the data input $\mathbf{X}$ train and $\mathbf{Y}$ train variables was visualized.
+The figures are showed in \ref{fig:yTrainClass} and \ref{fig:boxplotXClass}.
+After plotting this figures, all -1 values were changed on 0 in $\mathbf{Y}$ train and $\mathbf{X}$ train variables were normalized (normalizing the output variable allow us to obtain convergence faster with gradient decent and Newton's method).
+We tried gradient decent and Newton`s method for Logistic regression and Penalized Logistic Regression  and chose Newton`s method as the execution time was nearly the same, but Newton`s method was converged better than gradient descent method(as we needed to changing alpha parameter too) and with less iterations.
+
+Plotting histogram graphs for each variable in $\mathbf{X}$ train, showed that features have different distributions and mostly do not follow to normal Gaussian distribution.
+
+\begin{figure}[!t]
+\center
+\subfigure[Histogram of $\mathbf{Y}$ train.]{\includegraphics[width=2.0in]{figures/HistY.png} \label{fig:yTrainClass}}
+\subfigure[Boxplot of $\mathbf{X}$ train.]{\includegraphics[width=3.3in]{figures/boxplotX.png} \label{fig:boxplotXClass}}
+\hfill
+\caption{}
+\end{figure}
+
+\subsection{Penalized Logistic regression}
+We applied logistic regression and penalized logistic regression method for dataset.
+As we have many features, we expected that Penalized Logistic regression will improve test error result for kFold cross-validation with $k=4$ by penalizing some features.
+But the test 1-0 loss error for both methods was nearly the same and equals $0.183\pm0.001$.The summarized result with lambda in range between $10^{-4}$ to $10^3$  with 100 points presented in the Figure \ref{fig:pen}.
+
+\subsection{Feature transformation}
+Several features transformations were performed.
+Features deleting or using only classified features did not improve prediction result.
+Then we applied polynomial function transformation for each feature.
+We tried the same approach as with linear regression, we found features with two modes histogram distribution (the hypothesis is that features follow 0 and 1 output result) and use only them with polynomial feature transformation, but it did not improve the error result too.
+
+We compared penalized logistic regression without features engineering results with polynomial transformation.
+Method with transformation showed better result and gave the 1-0 loss test error result  $0.167\pm0.001$.
+The lambda parameter in case without feature transformation was chosen equal 0.001 and with feature transformation was chosen equal to 0.3 according to result of a k-fold cross validation procedure with $k=4$ and lambda in range between $10^{-3}$ to $10^4$  with 100 points.
+The result described here are summarized in the Figure \ref{fig:pen} and \ref{fig:poly_pen}.
+
+\begin{figure}[!t]
+\center
+\subfigure[Penalized logistic regression on kFold with $k=4$.]{\includegraphics[width=2.69in]{figures/pen_lambda.png} \label{fig:pen}}
+\subfigure[Penalized logistic regression on kFold with $k=4$ and feature feature transformation.]{\includegraphics[width=2.69in]{figures/pen_lambda_poly.png} \label{fig:poly_pen}}
+\hfill
+\caption{}
+\end{figure}
+
+\section{Summary}
+
+\section{Acknowledgments}
+We would like to thank Emtiyaz Khan and all teacher`s advisers for making the datasets for this project and answering our question during exercise sessions.
+
+The regression part was done by Corentin Tallec, as well as section Linear regression of report.
+The classification part was done by Roman Shirochenko, as well as the Classification part of the report.
+
 \end{document}

### 2015-11-01 report/paper.tex
Changed report tex file

@@
 \section{Summary}
+We analyzed a regression and classification datasets and that ridge regression and penalized logistic regression are reasonable fit.
+We tried some feature transformation and found that polynomial feature transformation gives us a little improvement in final test error result. 
+We estimate that the average test error for regression problem is $0.115\pm0.001$ on the test set, that is approximately one tenth of the variance of the output variable.
+The test error for classification problem is $0.167\pm0.001$.
 

### 2015-11-01 submit_files_folder/testCode.m
Added report to submit folder

@@ -1,62 +1,0 @@
-% Run this to make sure your implementation is correct
-% An example of writing csv file is given in the end of this file
-clear all;
-
-% generating dummy dataset. regression
-beta = [1,2,3,4,5]';
-N = 100;
-D = length(beta);
-tX = ones(N,D);
-for i = 2:D
-   tX(:,i) = randn(1,N) * i;
-end
-y = tX * beta;
-
-% max error
-eps = 1e-2;
-
-% test least squares via gradient descent
-disp('least squares via gradient descent...');
-alpha = 0.01;
-tBeta = leastSquaresGD(y, tX, alpha);
-assert(all(abs(tBeta - beta) < eps));
-disp('OK!');
-
-% testing least squares
-disp('least squares...');
-tBeta = leastSquares(y, tX);
-assert(all(abs(tBeta - beta) < eps));
-disp('OK!');
-
-% testing ridge regression
-disp('ridge regression...');
-lambda = 1e-5;
-tBeta = ridgeRegression(y, tX, lambda);
-assert(all(abs(tBeta - beta) < eps));
-disp('OK!');
-
-% generate binary data
-y = (y>0);
-
-% testing logistic regression
-disp('logistic regression...');
-alpha = 1e-3;
-tBeta = logisticRegression(y,tX,alpha);
-tY = 1.0 ./ (1.0 + exp(-tX * tBeta)) > 0.5;
-assert(sum(tY ~= y) / size(y,1) < 0.2);
-disp('OK!');
-
-% testing penalize logistic regression
-disp('penalized logistic regression...');
-alpha = 1e-3;
-lambda = 1e-2;
-tBeta = penLogisticRegression(y,tX,alpha,lambda);
-tY = 1.0 ./ (1.0 + exp(-tX * tBeta)) > 0.5;
-assert(sum(tY ~= y) / size(y,1) < 0.2);
-disp('OK!');
-
-% code for writing csv files
-tY = 1.0 ./ (1.0 + exp(-tX * tBeta));
-%csvwrite('predictions_classification.csv', tY);
-
-disp('test finished successfully');

### 2016-05-12 mlproject/report/paper.tex
Cleaning up the folder

--- 2015-11-01 report/paper.tex
@@
 
-\title{Project-I by Group CITYNAME}
+\title{Project-I by Group Rome}
 
@@
 \subsection{Data visualization and cleaning}
-Firstly, the data input $\mathbf{X}$ train and $\mathbf{Y}$ train variables was visualized.
-The figures are showed in \ref{fig:yTrainClass} and \ref{fig:boxplotXClass}.
-After plotting this figures, all -1 values were changed on 0 in $\mathbf{Y}$ train and $\mathbf{X}$ train variables were normalized (normalizing the output variable allow us to obtain convergence faster with gradient decent and Newton's method).
-We tried gradient decent and Newton`s method for Logistic regression and Penalized Logistic Regression  and chose Newton`s method as the execution time was nearly the same, but Newton`s method was converged better than gradient descent method(as we needed to changing alpha parameter too) and with less iterations.
+Firstly, the data input $\mathbf{X}$ train and $\mathbf{Y}$ train variables were
+visualized. The figures are showed in \ref{fig:yTrainClass} and
+\ref{fig:boxplotXClass}. After plotting these figures, all -1 values were
+changed to 0 in $\mathbf{Y}$ train and $\mathbf{X}$ train variables were
+normalized (normalizing the output variable gave us a faster convergence for
+both gradient decent and Newton's method). We tried gradient decent and Newton`s
+method for Logistic regression and Penalized Logistic Regression  and chose
+Newton`s method as the execution time was nearly the same, but Newton`s method
+displayed better convergence than gradient descent method (as we needed to changing alpha parameter too) and with less iterations.
 
-Plotting histogram graphs for each variable in $\mathbf{X}$ train, showed that features have different distributions and mostly do not follow to normal Gaussian distribution.
+Plotting histogram graphs for each variable in $\mathbf{X}$ train, showed that features have different distributions and mostly do not follow normal distribution.
 
@@
 \subsection{Penalized Logistic regression}
-We applied logistic regression and penalized logistic regression method for dataset.
-As we have many features, we expected that Penalized Logistic regression will improve test error result for kFold cross-validation with $k=4$ by penalizing some features.
-But the test 1-0 loss error for both methods was nearly the same and equals $0.183\pm0.001$.The summarized result with lambda in range between $10^{-4}$ to $10^3$  with 100 points presented in the Figure \ref{fig:pen}.
+We applied logistic regression and penalized logistic regression method on
+ourdataset. As we have many features and as the dataset is rank deficient, we expected that Penalized Logistic
+regression would improve test error result for kFold cross-validation with $k=4$
+by penalizing some features. But the test 1-0 loss error for both methods was
+nearly the same and equals to $0.183\pm0.001$. The summarized result with lambda
+ranging from $10^{-4}$ to $10^3$  with 100 points is presented in the Figure \ref{fig:pen}.
 
 \subsection{Feature transformation}
-Several features transformations were performed.
-Features deleting or using only classified features did not improve prediction result.
-Then we applied polynomial function transformation for each feature.
-We tried the same approach as with linear regression, we found features with two modes histogram distribution (the hypothesis is that features follow 0 and 1 output result) and use only them with polynomial feature transformation, but it did not improve the error result too.
+Several features transformations were performed. Features deleting or using only
+classified features did not improve prediction result. Then we applied
+polynomial function transformation for each feature. We tried the same approach
+as with linear regression, we found features with two modes histogram
+distribution (the hypothesis is that features follow 0 and 1 output result) and
+use only them with polynomial feature transformation, but it did not improve the
+error resuls.
 
-We compared penalized logistic regression without features engineering results with polynomial transformation.
-Method with transformation showed better result and gave the 1-0 loss test error result  $0.167\pm0.001$.
-The lambda parameter in case without feature transformation was chosen equal 0.001 and with feature transformation was chosen equal to 0.3 according to result of a k-fold cross validation procedure with $k=4$ and lambda in range between $10^{-3}$ to $10^4$  with 100 points.
-The result described here are summarized in the Figure \ref{fig:pen} and \ref{fig:poly_pen}.
+We compared penalized logistic regression without features engineering results
+with polynomial transformation. Method with transformation showed better result
+and gave the 1-0 loss test error result  $0.167\pm0.001$. The lambda parameter
+without feature transformation was chosen to be 0.001 and with feature
+transformation was chosen to be 0.3 according to the results of a k-fold cross
+validation procedure with $k=4$ and lambda ranging from $10^{-3}$ to $10^4$  with 100 points. The result described here are summarized in the Figure \ref{fig:pen} and \ref{fig:poly_pen}.
 
@@
 We tried some feature transformation and found that polynomial feature transformation gives us a little improvement in final test error result. 
-We estimate that the average test error for regression problem is $0.115\pm0.001$ on the test set, that is approximately one tenth of the variance of the output variable.
-The test error for classification problem is $0.167\pm0.001$.
+We estimate that the average test error for regression problem is $0.225\pm0.005$
+on the test set, that is approximately one fifth of the standard deviation of the output variable. The test error for classification problem is $0.167\pm0.001$.
+
+All the scripts that were used can be find on the webpage:
+\url{https:\\github.com/corentintallec/mlproject}
 

### 2016-05-26 coproject/report.tex
Added the report file.

\documentclass[a4paper, 11pt]{article}
\usepackage[english]{babel}

\title{Kernel Methods and SVMs:
A convex optimization point of view}
\author{Gaël Lederrey\\
  \and
Corentin Tallec}

\begin{document}
\maketitle
\begin{abstract}
  Kernel methods, and more specifically support vectors machines, are
  well known, well theorized and efficient machine learning tools. For
  years, they were presenting state of the art performances in many
  machine learning fields, before being overtaken by deep learning
  methods. They have the enormous advantage of being well understood
  theoretically, and most interestingly to belong to the class of
  convex optimization problems.

  In this document, we aim at presenting the general framework of
  Kernel methods, to expose how they relate to convex optimization,
  and how their optimization can be undertaken in practice. Besides,
  we intend to give a more precise treatment of the support vectors
  machines special case.
\end{abstract}
\section{Kernel methods}
\subsection{Generalities}
\subsection{Kernel methods as convex optimization problems}
\subsection{From infinite to finite}
\section{Support vector machines: a special case}
\subsection{Generalities}
\subsection{Dual problem}
\subsection{A use case}
\end{document}

### 2016-05-26 coproject/report.tex
Added the report file.

@@
   theoretically, and most interestingly to belong to the class of
-  convex optimization problems.
+  convex optimization problems, which is clearly not the case of the
+  latter.
 

### 2016-05-28 coproject/report_2.tex
ADD: report2.tex (avoid conflicts) and gitignore
ADD: Beginning of first part of the project

--- 2016-05-26 coproject/report.tex
@@
 
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+
+\newtheorem{theorem}{Theorem}
+
 \title{Kernel Methods and SVMs:
 A convex optimization point of view}
-\author{Gaël Lederrey\\
+\author{Gael Lederrey\\
   \and
@@
 \end{abstract}
+
 \section{Kernel methods}
 \subsection{Generalities}
+
+The kernel methods are used in machine learning for pattern analysis.
+The support vector machine (SVM) is the best know kernel method.
+This class of algorithm is used to find and study some types of relations in datasets, {\it e.g.} classifications of the data, correlations between them, etc.
+For a usual algorithm solving this kind of task, the data needs to be transformed into feature vector representations which can be computationally costly.
+The kernel methods require only a {\bf kernel}, {\it i.e.} a similarity function between pairs of data, specified by the user.
+\\
+In the following sections, we will first give the definition of a kernel function, then we will present the feature mapping and finally we will present the idea of the kernel method
+
+\subsubsection{Kernel}
+
+First, we denote $\mathcal{X}$ an arbitrary input set and $\mathcal{Y}$ an output set.
+An input-output pair $(x,y) \in \mathcal{X} \times \mathcal{Y}$ is called an example.
+A {\bf kernel} is a similarity function between these two sets.
+It's a real-valued function that quantifies the similarity between the input $x$ and the output $y$.
+We define the function
+$K: \mathcal{X}\times\mathcal{X} \mapsto \mathbb{R}$, a {\bf kernel} function if $K$ is symmetric and if it is positive semidefinite: for any $x_1, \ldots, x_m \in \mathcal{X}$, the Gram matrix $G\in \mathbb{R}^{m\times m}$, defined by $G_{ij} = K(x_i,x_j) = K(x_j,x_i) = G_{ji}$, is positive semidefinite.
+
+\subsubsection{Feature Mapping}
+
+In order to make the computation simpler, the kernel needs to be written in the form of a "feature map". Mercer's theorem (or Aronszajn's theorem) gives the definition of a feature map:
+\begin{theorem} [Mercer's theorem]
+The function $K$ is a positive definite kernel on the $\mathcal{X}$ if and only if there exists a Hilbert space $\mathcal{H}$ with an inner product $\langle \cdot,\cdot\rangle_{\mathcal{H}}$ and a mapping
+\[
+\phi: \mathcal{X}\mapsto\mathcal{H}
+\]
+such that, for any $x$, $x'$ in $\mathcal{X}$:
+\[
+K(x,x') = \langle\phi(x),\phi(x')\rangle_{\mathcal{H}}
+\]
+\end{theorem}
+The function $\phi$ is called the {\bf feature mapping}.
+\\
+In other words, the feature mapping is used to take the data in a space of small dimension $\mathcal{X}$ and map them into a Hilbert space of high dimension (even infinite).
+We do this because, it will be easier to perform a regression of the data in a space of higher dimension.
+
+\subsubsection{Idea about the Kernel Methods}
+\label{meth}
+The {\bf kernel methods} are using a set of pre-trained data.
+We can denote them by the pairs of example: $(x_1, y_1), \ldots, (x_n,y_n) \in \mathcal{X} \times \mathcal{Y}$.
+This means that for all the input $x_i, ~ i=1,\ldots,n$, we know the corresponding result $y_i$.
+Now, we can add an unlabelled value, let's say $x'$.
+The kernel function will be used to compare $x'$ with all the pre-trained inputs $x_i$.
+The kernel will then say if $x'$ is similar to $x_i$ or not.
+With adding weights for each of the pre-trained examples, an output $\hat{y}$ depending on the unlabelled input $x'$ can be computed.
+\\ 
+We can use the example of a {\it kernelized binary classifier}. It computes a weighted sum of similarities:
+\begin{equation}
+\hat{y} = \text{sgn}\sum_{i=1}^{n }w_i y_i K(x_i,x')
+\end{equation}
+where
+\begin{itemize}
+\item $\hat{y}\in\{-1,+1\}$ is the predicted output for the unlabeled input $x'$
+\item $K: \mathcal{X}\times\mathcal{X} \mapsto \mathbb{R}$ is the kernel function
+\item $\{(x_i, y_i)\}_{i=1}^{n}$ are the pre-trained examples with $y_i\in\{-1,+1\}\forall i$.
+\item $w_i\in \mathbb{R}$ are the weights for the pre-trained examples. They are determined by the larning algorithm.
+\item $\text{sgn}$ is the sign function.
+\end{itemize}
+With this example, we can clearly see how a kernel function can be used to label a new input. The kernel method is the learning algorithm that will provide the weights $w_i$.
+
 \subsection{Kernel methods as convex optimization problems}
-\subsection{From infinite to finite}
+
+From section \ref{meth}, we already understand that the learning algorithm will have to find the best weights in order to have the most accurate prediction.
+The idea of the learning algorithm is to take a subset of the pre-trained examples, test the predictions, change the weight, etc.
+This can easily be seen as an optimization problem.
+
+\subsection{From infinite to finite - The kernel trick}
 \section{Support vector machines: a special case}

### 2016-05-28 coproject/report_2.tex
UPD: Updated beginning of the report

@@
 \usepackage{amssymb}
+\usepackage{xcolor}
 
 \newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
 
@@
 \subsubsection{Feature Mapping}
+\label{map}
 
@@
 We do this because, it will be easier to perform a regression of the data in a space of higher dimension.
+
+{\color{red} Je devrais peut etre definir ici les RKHS}
 
@@
 This can easily be seen as an optimization problem.
+\\\\
 
-\subsection{From infinite to finite - The kernel trick}
+{\color{red} Je suis un peu perdu sur cette partie... }
+
+\subsection{From infinite to finite}
+
+{\bf \color{red} KERNEL TRICK + REPRESENTER THEOREM}
+
+We saw in the section \ref{map}, we can use a feature map to go from the input space $\mathcal{X}$ to a Hilbert space of a very high or infinite dimension $\mathcal{H}$.
+However, having the explicit representation of the mapping can be difficult and it is not convenient to work with it because of the high dimension of $\mathcal{H}$.
+Therefore, the idea is to work implicitly in the feature space $\mathcal{H}$.
+In order to do this, we must define the kernel trick.
+
+\begin{definition} [The Kernel trick]
+We take any algorithm to process finite-dimensional vectors.
+If we can express this algorithms only in terms of pairwise inner products.
+Then this algorithm can be applied to potentially infinite-dimensional vectors in the feature space of a positive definite kernel if we replace each inner product evaluation by a kernel evaluation.
+\end{definition}
+
+Intuitively, the idea is that we will compute $K(x,x') = \langle\phi(x),\phi(x')\rangle_{\mathcal{H}}$ without even knowing the function $\phi(x)$ because it is often really complex.
+\\\\
+
 \section{Support vector machines: a special case}

### 2016-05-28 coproject/report_2.tex
UPD: beginning of the report
Some corrections made, more to be added

@@
 \newtheorem{definition}{Definition}
+
+\newcommand{\X}{\mathcal{X}}
+\newcommand{\Y}{\mathcal{Y}}
+\newcommand{\F}{\mathcal{F}}
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\N}{\mathbb{N}}
+
+\newcommand{\x}{x_i}
+\newcommand{\y}{y_i}
+
+\newcommand{\inner}[2]{\langle #1 , #2 \rangle_{\R^d}}
+\newcommand{\norm}[1]{\left\| #1 \right\|}
 
@@
 
-The kernel methods are used in machine learning for pattern analysis.
-The support vector machine (SVM) is the best know kernel method.
-This class of algorithm is used to find and study some types of relations in datasets, {\it e.g.} classifications of the data, correlations between them, etc.
-For a usual algorithm solving this kind of task, the data needs to be transformed into feature vector representations which can be computationally costly.
-The kernel methods require only a {\bf kernel}, {\it i.e.} a similarity function between pairs of data, specified by the user.
-\\
-In the following sections, we will first give the definition of a kernel function, then we will present the feature mapping and finally we will present the idea of the kernel method
+Kernel methods are tools broadly used in various fields, but most notably in
+machine learning and pattern discrimination. Among this class of methods,
+support vector machines are probably the best known kernel method. This class
+of algorithm is used to find and study some types of relations in datasets,
+{\it e.g.} classifications of the data, correlations between them, etc. Many
+machine learning algorithms rely on the idea of embedding the data at hand
+into a vector space that makes it easier analyse. Kernel methods provide
+a quite general framework for doing so, and provide tools to transport data
+into possibly infinite dimensional spaces. Kernel methods rely on a measure
+of similarity between input data known as a {\bf kernel}, which has to be selected
+beforehand.
 
-\subsubsection{Kernel}
+A typical case of application for a kernel method is the following. Define
+$\X$ (resp. $\Y$) as a set of datapoints (resp. a set of labels). Given
+a training set, that is a set of couples $(\x, \y)_{i\leq n} \in (\X \times \Y)^n$,
+we want to find an $f : \X \mapsto \Y$ in a particular set of functions $\F$ such
+that $f$ predicts accurately labels in the training set, and is able to generalize
+to unseen datapoints. One way of doing so is to solve the following optimization
+problem:
+\begin{equation}
+  \min\limits_{f\in\F} \frac{1}{n}\sum\limits_{i=1}^nL(\y,f(\x)) + \lambda\Omega(f)
+\end{equation}
+where $L$ is a cost function, that measure how close $f(\x)$ is to $\y$ for each
+$i$, and $\Omega$ is a regularizing term, that aims at describing the complexity
+of a certain function $f$. In the general case, this optimization problem is not required
+to be convex. Kernel methods provide a set of function $\F$ and a regularizing term
+$\Omega(f)$ such that when the cost function is assumed to be convex, the entire
+problem is convex, and can be easily solved.
 
-First, we denote $\mathcal{X}$ an arbitrary input set and $\mathcal{Y}$ an output set.
-An input-output pair $(x,y) \in \mathcal{X} \times \mathcal{Y}$ is called an example.
-A {\bf kernel} is a similarity function between these two sets.
-It's a real-valued function that quantifies the similarity between the input $x$ and the output $y$.
-We define the function
-$K: \mathcal{X}\times\mathcal{X} \mapsto \mathbb{R}$, a {\bf kernel} function if $K$ is symmetric and if it is positive semidefinite: for any $x_1, \ldots, x_m \in \mathcal{X}$, the Gram matrix $G\in \mathbb{R}^{m\times m}$, defined by $G_{ij} = K(x_i,x_j) = K(x_j,x_i) = G_{ji}$, is positive semidefinite.
+In the following sections, kernel functions are defined, as well as their
+relation to the idea of feature mapping, and the general principle of kernel
+methods is exposed.
 
-\subsubsection{Feature Mapping}
+\subsubsection{Positive Definite Kernel}
+\label{subsub:psd}
+Denote by $\X$ an arbitrary input set (notably $\X$ is not required to be a vector
+space). A kernel can be seen as a measure of similarity between two elements of $\X$.
+Formally
+\begin{definition}
+  A positive definite (p.d.) kernel on $\X$ is a function $K:\X\times\X \mapsto \R$ that
+  is symmetric and that satisfies for all $N \in \N$, $(x_1, \dots, x_n) \in \X^N$,
+  $(a_1, \dots, a_N) \in \R^\N$:
+  \begin{equation}
+    \sum\limits_{1 \leq i, j \leq N}a_ia_jK(x_i,x_j) \geq 0.
+  \end{equation}
+\end{definition}
+If $\X = R^d$, the simplest kernel that can be thought of
+is the canonical inner product. It obviously is symmetric, 
+and positivity is easily verified. 
+Under the same assumptions, it can quite easily be shown
+that $K(x, x') = (\inner{x}{x'})^p$ is a kernel too (known
+as the polynomial kernel). Another well known kernel is 
+the gaussian kernel, defined as $K(x, x') = e^{- \frac{\norm{x-y}^2}{2\sigma^2}}$.
+
+\subsubsection{Feature mapping}
 \label{map}
+As seen in Subsection~\ref{subsub:psd}, the canonical inner
+product of a vector space is a kernel. Mercer's theorem provides
+a kind of reciprocal statement: any kernel can be viewed as an
+inner product in a certain hilbert space which is a functional
+space on $\mathcal{X}$.
+\begin{theorem} [Mercer's theorem]
+The function $K$ is a positive definite kernel on $\mathcal{X}$ if and only
+if there exists a Hilbert space $\mathcal{H}$ with an inner product $\langle
+\cdot,\cdot\rangle_{\mathcal{H}}$ and a mapping
+\begin{equation}
+  \phi: \mathcal{X}\mapsto\mathcal{H}
+\end{equation}
+such that, for any $x$, $x'$ in $\mathcal{X}$:
+\begin{equation}
+  K(x,x') = \langle\phi(x),\phi(x')\rangle_{\mathcal{H}}
+\end{equation}
+\end{theorem}
+Using a kernel can thus somehow be viewed as embedding the datapoints in
+a (larger) space. One benefit of doing so can be, for example, to simplify
+regression or classification problems, as these problems tend to
+be easier in high dimensional space than in low dimensional ones.
 
-In order to make the computation simpler, the kernel needs to be written in the form of a "feature map". Mercer's theorem (or Aronszajn's theorem) gives the definition of a feature map:
-\begin{theorem} [Mercer's theorem]
-The function $K$ is a positive definite kernel on the $\mathcal{X}$ if and only if there exists a Hilbert space $\mathcal{H}$ with an inner product $\langle \cdot,\cdot\rangle_{\mathcal{H}}$ and a mapping
-\[
-\phi: \mathcal{X}\mapsto\mathcal{H}
-\]
-such that, for any $x$, $x'$ in $\mathcal{X}$:
-\[
-K(x,x') = \langle\phi(x),\phi(x')\rangle_{\mathcal{H}}
-\]
-\end{theorem}
-The function $\phi$ is called the {\bf feature mapping}.
-\\
-In other words, the feature mapping is used to take the data in a space of small dimension $\mathcal{X}$ and map them into a Hilbert space of high dimension (even infinite).
-We do this because, it will be easier to perform a regression of the data in a space of higher dimension.
-
-{\color{red} Je devrais peut etre definir ici les RKHS}
-
+\subsubsection{Back to classification}
 \subsubsection{Idea about the Kernel Methods}

### 2016-05-28 coproject/report_2.tex
Finished corrections. I think the last two subsubsections can be removed. SVMs to be done by tomorrow.

@@
 \newcommand{\N}{\mathbb{N}}
+\newcommand{\Hilb}{\mathcal{H}}
 
@@
 \newcommand{\inner}[2]{\langle #1 , #2 \rangle_{\R^d}}
+\newcommand{\innerh}[2]{\langle #1 , #2 \rangle_{\Hilb}}
 \newcommand{\norm}[1]{\left\| #1 \right\|}
+\newcommand{\normh}[1]{\left\| #1 \right\|_\Hilb}
 
@@
   machines special case.
+
+  \color{red}We need to add the reference to the course notes we
+  used !
+  \color{black}
 \end{abstract}
@@
 \subsection{Generalities}
+\label{sub:gen}
 
@@
   \min\limits_{f\in\F} \frac{1}{n}\sum\limits_{i=1}^nL(\y,f(\x)) + \lambda\Omega(f)
+  \label{eq:min}
 \end{equation}
-where $L$ is a cost function, that measure how close $f(\x)$ is to $\y$ for each
-$i$, and $\Omega$ is a regularizing term, that aims at describing the complexity
-of a certain function $f$. In the general case, this optimization problem is not required
-to be convex. Kernel methods provide a set of function $\F$ and a regularizing term
-$\Omega(f)$ such that when the cost function is assumed to be convex, the entire
-problem is convex, and can be easily solved.
+where $L$ is a cost function, that measure how close $f(\x)$ is to $\y$ for
+each $i$, $\lambda$ is a positive real number and $\Omega$ is a regularizing
+term, that aims at describing the complexity of a certain function $f$. In the
+general case, this optimization problem is not required to be convex. Kernel
+methods provide a set of function $\F$ and a regularizing term $\Omega(f)$ such
+that when the cost function is assumed to be convex, the entire problem is
+convex, and can be easily solved.
 
@@
 \subsubsection{Feature mapping}
-\label{map}
+\label{subsub:map}
 As seen in Subsection~\ref{subsub:psd}, the canonical inner
@@
 
-\subsubsection{Back to classification}
+\subsubsection{Kernel methods as convex optimization problems}
+Recall the kind of problem exposed in Sec~\ref{sub:gen}, by eq~\ref{eq:min}.
+Now, once a kernel $K$ is defined, it implicitly defines a functional space $\Hilb$ on $\X$, as
+explained in Sec~\ref{subsub:map}. This space can be used as the prediction function space,
+that is $\F = \Hilb$, and the regularization term can be replaced by the squared norm in $\Hilb$.
+This gives the following minimization problem:
+\begin{equation}
+  \min\limits_{f\in\Hilb} \frac{1}{n}\sum\limits_{i=1}^nL(\y,f(\x)) + \lambda\normh{f}^2
+  \label{eq:minkernel}
+\end{equation}
+
+When $L$ is a convex function, it is quite clear that the resulting minimization problem is
+a convex optimization problem: the squared norm is a convex function of $f$, $f \mapsto f(x)$
+is a linear function for any $x \in \X$, thus $L(\y,f(\x))$ is convex for all $i$, which lead
+to the global convexity. However, the space $\Hilb$ considered is, in the general case, infinite,
+and this makes convex optimization unapplicable.
+
+\subsection{From infinite to finite}
+Happily enough, the representer theorem stated below turns this infinite dimensional convex optimization
+problem into your gentle everyday $n$-dimensional convex optimization problem:
+\begin{theorem}[Representer theorem]
+  Using the notations previously defined, let $\Psi:\R^{n+1}\mapsto\R$ be a function strictly increasing in 
+  the last variable. Then any solution of
+  \begin{equation}
+    \min\limits_{f\in\Hilb}\Psi(f(x_1), \dots, f(x_n), \normh{f})
+  \end{equation}
+  admits a representation of the form:
+  \begin{equation}
+    \forall x \in \X, f(x) = \sum\limits_{i=1}^n\alpha_iK(\x,x).
+  \end{equation}
+\end{theorem}
+The equation~\ref{eq:minkernel} clearly falls in the field of application of the representer theorem.
+The optimization problem can thus be rewritten in term of $\alpha$'s as:
+\begin{equation}
+  \min\limits_{(\alpha_1, \dots, \alpha_n)\in\R^n} 
+  \frac{1}{n}\sum\limits_{i=1}^nL(\y, \sum\limits_{j=1}^n\alpha_jK(x_j, x_i)) +
+  \lambda\sum\limits_{1\leq i,j\leq n}\alpha_i\alpha_j K(\x, x_j)
+\end{equation}
+which is a finite dimensional convex optimization problem.
+
 \subsubsection{Idea about the Kernel Methods}
-\label{meth}
+\label{subsub:meth}
 The {\bf kernel methods} are using a set of pre-trained data.
@@
 
-From section \ref{meth}, we already understand that the learning algorithm will have to find the best weights in order to have the most accurate prediction.
+From section, we already understand that the learning algorithm will have to find the best weights in order to have the most accurate prediction.
 The idea of the learning algorithm is to take a subset of the pre-trained examples, test the predictions, change the weight, etc.
@@
 
-We saw in the section \ref{map}, we can use a feature map to go from the input space $\mathcal{X}$ to a Hilbert space of a very high or infinite dimension $\mathcal{H}$.
+We saw in the section , we can use a feature map to go from the input space $\mathcal{X}$ to a Hilbert space of a very high or infinite dimension $\mathcal{H}$.
 However, having the explicit representation of the mapping can be difficult and it is not convenient to work with it because of the high dimension of $\mathcal{H}$.

### 2016-05-29 coproject/matlab/loadImages.m
ADD: matlab folder for svm

@@ -1,0 +1,25 @@
+function images = loadImages(fileName)
+% This function is used to load the images given a file
+
+file = fopen(fileName, 'rb');
+
+magicNumber = fread(file, 1, 'int32', 0, 'ieee-be');
+assert(magicNumber == 2051, ['Bad magic number in ', fileName, '']);
+
+numImages = fread(file, 1, 'int32', 0, 'ieee-be');
+numRows = fread(file, 1, 'int32', 0, 'ieee-be');
+numCols = fread(file, 1, 'int32', 0, 'ieee-be');
+
+images = fread(file, inf, 'unsigned char');
+images = reshape(images, numCols, numRows, numImages);
+images = permute(images,[2 1 3]);
+
+fclose(file);
+
+% Reshape to #pixels x #examples
+images = reshape(images, size(images, 1) * size(images, 2), size(images, 3));
+% Convert to double and rescale to [0,1]
+images = double(images) / 255;
+
+end
+

### 2016-05-29 coproject/matlab/loadLabels.m

@@ -1,0 +1,16 @@
+function labels = loadLabels(fileName)
+% This function is used to load the labels given a file
+
+file = fopen(fileName, 'rb');
+
+magicNumber = fread(file, 1, 'int32', 0, 'ieee-be');
+assert(magicNumber == 2049, ['Bad magic number in ', fileName, '']);
+
+numLabels = fread(file, 1, 'int32', 0, 'ieee-be');
+
+labels = fread(file, inf, 'unsigned char');
+
+fclose(file);
+
+end
+

### 2016-05-29 coproject/matlab/svm.m

@@ -1,0 +1,58 @@
+% load training set and testing set
+clear all;
+trainSet = loadImages('./data/train-images.idx3-ubyte')';
+trainLabel = strtrim(cellstr(num2str(loadLabels('./data/train-labels.idx1-ubyte'))));
+testSet = loadImages('./data/t10k-images.idx3-ubyte')';
+testLabel = strtrim(cellstr(num2str(loadLabels('./data/t10k-labels.idx1-ubyte'))));
+
+SVMModelsLin = cell(9,1);
+SVMModelsPol = cell(9,1);
+SVMModelsRBF = cell(9,1);
+classes = unique(trainLabel);
+
+rng(1); % For reproductibility
+
+trainingTimeLin = 0;
+trainingTimePol = 0;
+trainingTimeRBF = 0;
+
+times = zeros(10,3);
+
+%% TRAINING
+
+for i = 1:numel(classes)
+    display(['Start training for class ', num2str(i-1)]);
+    idx = strcmp(trainLabel, classes(i)); % Create binary classes for each classifier
+    
+    display(['Start Linear Kernel']);
+    tic;
+    SVMModelsLin{i} = fitcsvm(trainSet , idx , 'KernelFunction', 'linear');
+    times(i,1) = toc;
+    display(['Finished training for Linear Kernel, Time taken: ', num2str(times(i))]);
+    trainingTimeLin = trainingTimeLin + times(i,1);
+    
+    display(['Start Polynomial Kernel']);
+    tic;
+    SVMModelsLin{i} = fitcsvm(trainSet , idx , 'KernelFunction', 'polynomial', 'PolynomialOrder', 4);
+    times(i,1) = toc;
+    display(['Finished training for Polynomial Kernel, Time taken: ', num2str(times(i))]);
+    trainingTimeLin = trainingTimeLin + times(i,2);
+    
+    display(['Start RBF Kernel']);
+    tic;
+    SVMModelsLin{i} = fitcsvm(trainSet , idx , 'KernelFunction', 'rbf');
+    times(i,1) = toc;
+    display(['Finished training for RBF Kernel, Time taken: ', num2str(times(i))]);
+    trainingTimeLin = trainingTimeLin + times(i,3);    
+end
+
+%% CLASSIFICATION
+
+
+% classification
+%tic;
+%[PredictedLabel, accuracy, decisionValues] = svmpredict(testSet, testLabel, model);
+%t2 = toc;
+%hel
+%disp(num2str(t1));
+%disp(num2str(t2));

### 2016-05-29 coproject/matlab/svm-train.m
UPD: Matlab code

@@ -1,9 +1,9 @@
-% load training set and testing set
+%% LOADING DATA
 clear all;
 trainSet = loadImages('./data/train-images.idx3-ubyte')';
 trainLabel = strtrim(cellstr(num2str(loadLabels('./data/train-labels.idx1-ubyte'))));
-testSet = loadImages('./data/t10k-images.idx3-ubyte')';
-testLabel = strtrim(cellstr(num2str(loadLabels('./data/t10k-labels.idx1-ubyte'))));
+%testSet = loadImages('./data/t10k-images.idx3-ubyte')';
+%testLabel = strtrim(cellstr(num2str(loadLabels('./data/t10k-labels.idx1-ubyte'))));
 
 SVMModelsLin = cell(9,1);
 SVMModelsPol = cell(9,1);
@@ -30,21 +30,41 @@ for i = 1:numel(classes)
     times(i,1) = toc;
     display(['Finished training for Linear Kernel, Time taken: ', num2str(times(i))]);
     trainingTimeLin = trainingTimeLin + times(i,1);
-    
+end
+
+save('SVMModelsLin.mat', SVMModelsLin);
+save('trainingTimeLin.mat', trainingTimeLin);
+ 
+for i = 1:numel(classes)
+    display(['Start training for class ', num2str(i-1)]);
+    idx = strcmp(trainLabel, classes(i)); % Create binary classes for each classifier
+       
     display(['Start Polynomial Kernel']);
     tic;
-    SVMModelsLin{i} = fitcsvm(trainSet , idx , 'KernelFunction', 'polynomial', 'PolynomialOrder', 4);
-    times(i,1) = toc;
+    SVMModelsPol{i} = fitcsvm(trainSet , idx , 'KernelFunction', 'polynomial', 'PolynomialOrder', 4);
+    times(i,2) = toc;
     display(['Finished training for Polynomial Kernel, Time taken: ', num2str(times(i))]);
-    trainingTimeLin = trainingTimeLin + times(i,2);
+    trainingTimePol = trainingTimePol + times(i,2);
+end
+
+save('SVMModelsPol.mat', SVMModelsPol);
+save('trainingTimePol.mat', trainingTimePol);
+
+for i = 1:numel(classes)
+    display(['Start training for class ', num2str(i-1)]);
+    idx = strcmp(trainLabel, classes(i)); % Create binary classes for each classifier
     
[…]

### 2016-05-29 coproject/matlab/svm-train.m
UPD: Matlab code for training

@@ -20,45 +20,50 @@ times = zeros(10,3);
 
 %% TRAINING
 
+display('Start training for Linear Kernel');
+
 for i = 1:numel(classes)
-    display(['Start training for class ', num2str(i-1)]);
+    display(['Start training for digit ', num2str(i-1)]);
     idx = strcmp(trainLabel, classes(i)); % Create binary classes for each classifier
     
     display(['Start Linear Kernel']);
     tic;
     SVMModelsLin{i} = fitcsvm(trainSet , idx , 'KernelFunction', 'linear');
     times(i,1) = toc;
-    display(['Finished training for Linear Kernel, Time taken: ', num2str(times(i))]);
+    display(['Finished training for digit ', num2str(i-1), '; Time taken: ', num2str(times(i))]);
     trainingTimeLin = trainingTimeLin + times(i,1);
 end
 
 save('SVMModelsLin.mat', SVMModelsLin);
 save('trainingTimeLin.mat', trainingTimeLin);
+
+display('Start training for Polynomial Kernel');
  
 for i = 1:numel(classes)
-    display(['Start training for class ', num2str(i-1)]);
+    display(['Start training for digit ', num2str(i-1)]);
     idx = strcmp(trainLabel, classes(i)); % Create binary classes for each classifier
-       
-    display(['Start Polynomial Kernel']);
+  
     tic;
     SVMModelsPol{i} = fitcsvm(trainSet , idx , 'KernelFunction', 'polynomial', 'PolynomialOrder', 4);
     times(i,2) = toc;
-    display(['Finished training for Polynomial Kernel, Time taken: ', num2str(times(i))]);
+    display(['Finished training for digit ', num2str(i-1), '; Time taken: ', num2str(times(i))]);
     trainingTimePol = trainingTimePol + times(i,2);
 end
 
 save('SVMModelsPol.mat', SVMModelsPol);
 save('trainingTimePol.mat', trainingTimePol);
 
+display('Start training for RBF Kernel');
+
 for i = 1:numel(classes)
-    display(['Start training for class ', num2str(i-1)]);
+    display(['Start training for digit ', num2str(i-1)]);
     idx = strcmp(trainLabel, classes(i)); % Create binary classes for each classifier
     
     display(['Start RBF Kernel']);
     tic;
     SVMModelsRBF{i} = fitcsvm(trainSet , idx , 'KernelFunction', 'rbf');
     times(i,3) = toc;
[…]

### 2016-05-29 coproject/matlab/train.m
Changed name of the train file.

@@ -1,5 +1,8 @@
 %% LOADING DATA
 clear all;
+close all;
+clc;
+
 trainSet = loadImages('./data/train-images.idx3-ubyte')';
 trainLabel = strtrim(cellstr(num2str(loadLabels('./data/train-labels.idx1-ubyte'))));
 %testSet = loadImages('./data/t10k-images.idx3-ubyte')';
@@ -26,7 +29,6 @@ for i = 1:numel(classes)
     display(['Start training for digit ', num2str(i-1)]);
     idx = strcmp(trainLabel, classes(i)); % Create binary classes for each classifier
     
-    display(['Start Linear Kernel']);
     tic;
     SVMModelsLin{i} = fitcsvm(trainSet , idx , 'KernelFunction', 'linear');
     times(i,1) = toc;
@@ -59,7 +61,6 @@ for i = 1:numel(classes)
     display(['Start training for digit ', num2str(i-1)]);
     idx = strcmp(trainLabel, classes(i)); % Create binary classes for each classifier
     
-    display(['Start RBF Kernel']);
     tic;
     SVMModelsRBF{i} = fitcsvm(trainSet , idx , 'KernelFunction', 'rbf');
     times(i,3) = toc;

### 2016-05-29 coproject/matlab/train.m
UPD: Add classify file

@@ -8,9 +8,9 @@ trainLabel = strtrim(cellstr(num2str(loadLabels('./data/train-labels.idx1-ubyte'
 %testSet = loadImages('./data/t10k-images.idx3-ubyte')';
 %testLabel = strtrim(cellstr(num2str(loadLabels('./data/t10k-labels.idx1-ubyte'))));
 
-SVMModelsLin = cell(9,1);
-SVMModelsPol = cell(9,1);
-SVMModelsRBF = cell(9,1);
+SVMModelsLin = cell(10,1);
+SVMModelsPol = cell(10,1);
+SVMModelsRBF = cell(10,1);
 classes = unique(trainLabel);
 
 rng(1); % For reproductibility
@@ -21,7 +21,7 @@ trainingTimeRBF = 0;
 
 times = zeros(10,3);
 
-%% TRAINING
+%% TRAINING LIN
 
 display('Start training for Linear Kernel');
 
@@ -36,10 +36,14 @@ for i = 1:numel(classes)
     trainingTimeLin = trainingTimeLin + times(i,1);
 end
 
-save('SVMModelsLin.mat', SVMModelsLin);
-save('trainingTimeLin.mat', trainingTimeLin);
+save('SVMModelsLin', 'SVMModelsLin','-v7.3');
+save('times', 'times');
+
+%% TRAINING POL
 
 display('Start training for Polynomial Kernel');
+
+load('times');
  
 for i = 1:numel(classes)
     display(['Start training for digit ', num2str(i-1)]);
@@ -52,10 +56,14 @@ for i = 1:numel(classes)
     trainingTimePol = trainingTimePol + times(i,2);
 end
 
-save('SVMModelsPol.mat', SVMModelsPol);
-save('trainingTimePol.mat', trainingTimePol);
+save('SVMModelsPol', 'SVMModelsPol','-v7.3');
+save('times', 'times');
+
+%% TRAINING RBF
 
 display('Start training for RBF Kernel');
+
+load('times');
 
 for i = 1:numel(classes)
     display(['Start training for digit ', num2str(i-1)]);
@@ -68,17 +76,5 @@ for i = 1:numel(classes)
     trainingTimeRBF = trainingTimeRBF + times(i,3);    
 end
 
-save('SVMModelsRBF.mat', SVMModelsRBF);
-save('trainingTimeRBF.mat', trainingTimeRBF);
-save('times.mat',times);
-
-%% CLASSIFICATION
-
[…]

### 2016-05-29 coproject/report_2.tex
Text finished. To be reviewed.

@@
 \usepackage{xcolor}
+\usepackage{hyperref}
 
@@
 \newcommand{\N}{\mathbb{N}}
+\newcommand{\K}{\bold{K}}
 \newcommand{\Hilb}{\mathcal{H}}
@@
 \newcommand{\y}{y_i}
+\newcommand{\pmset}{\left\{\pm1\right\}}
 
@@
 
-  \color{red}We need to add the reference to the course notes we
-  used !
-  \color{black}
+  Most theoretical results used here are presented, explained and
+  proved more thoroughly in the ``Master vision et apprentissage''
+  of the ``Ecole normale superieure de Cachan'' course on kernels.
 \end{abstract}
@@
 methods provide a set of function $\F$ and a regularizing term $\Omega(f)$ such
-that when the cost function is assumed to be convex, the entire problem is
+that when the cost function is assumed to be convex in its second argument, the entire problem is
 convex, and can be easily solved.
@@
 that is $\F = \Hilb$, and the regularization term can be replaced by the squared norm in $\Hilb$.
+
 This gives the following minimization problem:
 \begin{equation}
-  \min\limits_{f\in\Hilb} \frac{1}{n}\sum\limits_{i=1}^nL(\y,f(\x)) + \lambda\normh{f}^2
+  \min\limits_{f\in\Hilb} \frac{1}{n}\sum\limits_{i=1}^nL(\y,f(\x)) + \lambda\normh{f}^2.
   \label{eq:minkernel}
 \end{equation}
-
-When $L$ is a convex function, it is quite clear that the resulting minimization problem is
+When $L$ is a convex function over its second argument, it is quite clear that the resulting minimization problem is
 a convex optimization problem: the squared norm is a convex function of $f$, $f \mapsto f(x)$
@@
     \forall x \in \X, f(x) = \sum\limits_{i=1}^n\alpha_iK(\x,x).
+    \label{eq:bestfit}
   \end{equation}
@@
 
-\subsubsection{Idea about the Kernel Methods}
-\label{subsub:meth}
-The {\bf kernel methods} are using a set of pre-trained data.
-We can denote them by the pairs of example: $(x_1, y_1), \ldots, (x_n,y_n) \in \mathcal{X} \times \mathcal{Y}$.
-This means that for all the input $x_i, ~ i=1,\ldots,n$, we know the corresponding result $y_i$.
-Now, we can add an unlabelled value, let's say $x'$.
-The kernel function will be used to compare $x'$ with all the pre-trained inputs $x_i$.
-The kernel will then say if $x'$ is similar to $x_i$ or not.
-With adding weights for each of the pre-trained examples, an output $\hat{y}$ depending on the unlabelled input $x'$ can be computed.
-\\ 
-We can use the example of a {\it kernelized binary classifier}. It computes a weighted sum of similarities:
+\section{Support vector machines: a special case}
+Support vector machines are kernel methods used in the field of binary
+classification, and that can be extended (quite painfully, unfortunately) to
+$n$-ary classification.  Among kernel methods, they are probably the best
+known.  This might be explained by the fact that they enforce sparsity in the
+best fit representation. Formally, they ensure that in eq~\ref{eq:bestfit}, a
+varying number of $\alpha$'s will be exactly zero.  Sparsity often brings along
+a lot of nice properties. For SVMs, most notably, they ensure a better
+regularization (intuitively, a fit with less term is less complex than a fit
+with many terms, and the number of term is directly related to the number of
+non-zero $\alpha$'s), and they induce smaller computational cost. Indeed, the
+best fit represented in eq~\ref{eq:bestfit} has as many terms as there are
+datapoints in the train dataset. This means that the computational cost at test
+time will depend directly on the number of datapoints in the training set. This
+might prove computationally costly. SVMs partly solve this problem by ensuring
+that some (most) $\alpha$'s are zero, and thus that the best fit will only
+depend on a few training datapoints. 
+\subsection{Generalities}
+Given an input set $\X$, a p.s.d kernel on $\X$, $K$ and a training dataset,
+$((x_1, y_1), \dots, (x_n, y_n)) \in (\X \times \pmset)^n$, SVMs solve the
+kernel optimization problem eq~\ref{eq:minkernel} with the particular choice of
+loss $L(x, y) = \max(0, 1 - x^\top y)$, which is known as the Hinge Loss.
+
+A few remarks might be of use at this point:
+\begin{itemize}
+  \item The Hinge Loss is clearly convex in $x$. It is a maximum over two affine
+    functions of $x$, and is thus convex.
+  \item Datapoints are classified simply by looking at the sign of $f(x)$ where
+    $f$ is the fit. If $f(x) > 0$, $x$ is classified as $+1$, else $x$ is classified
+    ad $-1$. The specificity of the Hinge Loss is to try and drive $f(x)$ towards $+1$
+    when $x$ is in the $+1$ class, and toward $-1$ when $x$ is in the $-1$ class, but not
+    to do any further effort.
+\end{itemize}
+Rewritting eq~\ref{eq:minkernel} in the SVM case, and introducing
+additionnal slack variables, the optimization problem we intend to
+solve takes the form:
 \begin{equation}
-\hat{y} = \text{sgn}\sum_{i=1}^{n }w_i y_i K(x_i,x')
+  \min\limits_{f\in\Hilb,\xi\in\R^n}
+  \frac{1}{n}\sum\limits_{i=1}^n\xi_i + \lambda \normh{f}^2
 \end{equation}
-where
-\begin{itemize}
-\item $\hat{y}\in\{-1,+1\}$ is the predicted output for the unlabeled input $x'$
-\item $K: \mathcal{X}\times\mathcal{X} \mapsto \mathbb{R}$ is the kernel function
-\item $\{(x_i, y_i)\}_{i=1}^{n}$ are the pre-trained examples with $y_i\in\{-1,+1\}\forall i$.
-\item $w_i\in \mathbb{R}$ are the weights for the pre-trained examples. They are determined by the larning algorithm.
-\item $\text{sgn}$ is the sign function.
-\end{itemize}
-With this example, we can clearly see how a kernel function can be used to label a new input. The kernel method is the learning algorithm that will provide the weights $w_i$.
+subject to:
+\begin{equation}
+  \begin{cases}
+    \xi_i \geq 1 - y_i f(x_i)\\
+    \xi_i \geq 0
+  \end{cases}
+\end{equation}
+Using the representer theorem, this can be transformed in
+\begin{equation}
+  \min_{\alpha\in\R^n, \xi\in\R^n}
+  \frac{1}{n}\sum\limits_{i=1}^n\xi_i
+  + \lambda \alpha^\top \K \alpha
+\end{equation}
+subject to:
+\begin{equation}
+  \begin{cases}
+    y_i\sum\limits_{j=1}^n\alpha_jK(\x, x_j) + \xi_i - 1 \geq 0\\
+    \xi_i \geq 0
+  \end{cases}
+\end{equation}
+where $\K$ is the p.s.d. matrix associated to the kernel $K$ on
+the dataset, that is $\K_{i,j} = K(\x, x_j)$. This is a quadratic
+program.
 
-\subsection{Kernel methods as convex optimization problems}
+The dual problem can easily be expressed:
+\begin{equation}
+  \max\limits_{0\leq\mu\leq 1/n}
+  \sum\limits_{i=1}^n\mu_i
+  - \frac{1}{4\lambda}\sum\limits_{1\leq i,j \leq n}y_iy_j\mu_i\mu_jK(x_i, x_j).
+\end{equation}
+Now expressing the relation between primal and dual solutions:
+\begin{equation}
+  \alpha = \text{diag}(y)\mu/2\lambda
+\end{equation}
+(this is done by rewriting the optimization condition on the
+Lagrangian with respect to the $\alpha$ variables)
+and injecting this relation into the dual, we get the following
+easier problem:
+\begin{equation}
+  2\alpha^\top y - \alpha^\top \K \alpha
+\end{equation}
+subject to:
+\begin{equation}
+  0 \leq y_i\alpha_i \leq \frac{1}{2\lambda n}
+\end{equation}
+(The crucial point in deriving this new optimization problem is
+to notice that $y_i^2 = 1$, since $y_i \in \pmset$).
 
-From section, we already understand that the learning algorithm will have to find the best weights in order to have the most accurate prediction.
-The idea of the learning algorithm is to take a subset of the pre-trained examples, test the predictions, change the weight, etc.
-This can easily be seen as an optimization problem.
-\\\\
-
-{\color{red} Je suis un peu perdu sur cette partie... }
-
-\subsection{From infinite to finite}
-
-{\bf \color{red} KERNEL TRICK + REPRESENTER THEOREM}
-
-We saw in the section , we can use a feature map to go from the input space $\mathcal{X}$ to a Hilbert space of a very high or infinite dimension $\mathcal{H}$.
-However, having the explicit representation of the mapping can be difficult and it is not convenient to work with it because of the high dimension of $\mathcal{H}$.
-Therefore, the idea is to work implicitly in the feature space $\mathcal{H}$.
-In order to do this, we must define the kernel trick.
-
-\begin{definition} [The Kernel trick]
-We take any algorithm to process finite-dimensional vectors.
-If we can express this algorithms only in terms of pairwise inner products.
-Then this algorithm can be applied to potentially infinite-dimensional vectors in the feature space of a positive definite kernel if we replace each inner product evaluation by a kernel evaluation.
-\end{definition}
-
-Intuitively, the idea is that we will compute $K(x,x') = \langle\phi(x),\phi(x')\rangle_{\mathcal{H}}$ without even knowing the function $\phi(x)$ because it is often really complex.
-\\\\
-
-\section{Support vector machines: a special case}
-\subsection{Generalities}
-\subsection{Dual problem}
-\subsection{A use case}
+On this very simple optimization problem, the KKT conditions can
+be easily expressed, and they lead to the following complementary
+slackness equations:
+\begin{equation}
+  \begin{cases}
+    \alpha_i\left[y_if(x_i) + \xi_i - 1\right] = 0\\
+    \left[\alpha_i - \frac{y_i}{2\lambda n}\right]\xi_i = 0.
+  \end{cases}
+\end{equation}
+With a bit of analysis with these equations and the constraints
+of the primal, it can be easily shown that the only terms for
+which $\alpha_i \neq 0$ are the examples ``hard to classify'' that
+is for which $y_i f(x_i) \leq 1$ (those examples can be well
+classified, but their classifiction margin is not large). All in
+all, this tends to express the fact that only a few datapoints
+are involved in the best fit expression. This has a tremendous
+impact on further computations involving the best fit.
 \end{document}

### 2016-06-06 coproject/code/svm.py
UPD: Changed the code for the project. Use python instead of matlab

@@ -1,0 +1,86 @@
+#!/usr/bin/python
+
+import matplotlib.pyplot as plt
+import numpy as np
+# Import datasets, classifiers and performance metrics
+from sklearn import datasets, svm, metrics
+from sklearn.cross_validation import StratifiedShuffleSplit
+from sklearn.grid_search import GridSearchCV
+from sklearn.svm import SVC
+import argparse
+
+from mnist import MNIST
+
+DATA_PATH='./data'
+SIZE_TRAIN=60000
+SIZE_TEST=10000
+
+unique_labels=[0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
+
+def main(kernel):
+  print "Loading the data"
+
+  mn = MNIST(DATA_PATH)
+
+  test_img, test_label = mn.load_testing()
+  train_img, train_label = mn.load_training()
+  
+  train_img = np.array(train_img[:SIZE_TRAIN])
+  train_label = np.array(train_label[:SIZE_TRAIN])
+  test_img = np.array(test_img[:SIZE_TEST])
+  test_label = np.array(test_label[:SIZE_TEST])
+
+  print "Finished loading the data"
+
+  # Create a classifier: a support vector classifier
+  if kernel == 'rbf':
+    print "Training with RBF kernel - Might take a few minutes"
+    classifier = svm.SVC(C=10, gamma=5e-7, kernel='rbf') 
+  elif kernel == 'linear':
+    print "Training with Linear kernel - Might take a few minutes"
+    classifier = svm.SVC(C=1e-6, kernel='linear')
+  elif kernel == 'poly':
+    print "Training with Polynomial kernel - Might take a few minutes"
+    #classifier = svm.SVC(C=10, gamma=1e-7, kernel='poly', degree=2)
+    #classifier = svm.SVC(C=10, gamma=1e-6, kernel='poly', degree=3)
+    classifier = svm.SVC(C=10, gamma=1e-6, kernel='poly', degree=4)
+
+  # We learn the digits on the first half of the digits
+  classifier.fit(train_img, train_label)
+
+  print "Classifying - Might take a few minutes"
+
+  predicted = classifier.predict(test_img)
+  print predicted
+
+  cm = metrics.confusion_matrix(test_label, predicted)
+
+  print("Classification report for classifier %s:\n%s\n"% (classifier, metrics.classification_report(test_label, predicted)))
+  print("Confusion matrix:\n%s" % cm)
+
[…]

### 2016-06-06 coproject/matlab/classify.m

@@ -1,45 +1,0 @@
-%% LOADING DATA
-
-clear all;
-close all;
-clc;
-
-testSet = loadImages('./data/t10k-images.idx3-ubyte')';
-testLabel = strtrim(cellstr(num2str(loadLabels('./data/t10k-labels.idx1-ubyte'))));
-testLabelNum = loadLabels('./data/t10k-labels.idx1-ubyte');
-
-% Change the model here
-load('SVMModelsLin');
-load('times');
-
-classes = unique(testLabel);
-
-%% PREDICT FOR LIN
-
-prediction = zeros(10,10);
-
-for i=1:numel(classes)
-    display(['Classify for digit ', num2str(i-1)]);
-    
-    predictedLabels = predict(SVMModelsLin{i}, testSet);
-    for j=1:length(predictedLabels)
-        if predictedLabels(j) == 1        
-           prediction(i, testLabelNum(j)+1) =  prediction(i,testLabelNum(j)+1) + 1;
-        end
-
-    end
-    
-    nbr = sum(predictedLabels);
-    
-    prediction(i,:) = prediction(i,:)./nbr*100;
-end
-
-prediction
-
-
-
-
-
-
-
-

### 2016-06-06 coproject/matlab/loadImages.m

@@ -1,25 +1,0 @@
-function images = loadImages(fileName)
-% This function is used to load the images given a file
-
-file = fopen(fileName, 'rb');
-
-magicNumber = fread(file, 1, 'int32', 0, 'ieee-be');
-assert(magicNumber == 2051, ['Bad magic number in ', fileName, '']);
-
-numImages = fread(file, 1, 'int32', 0, 'ieee-be');
-numRows = fread(file, 1, 'int32', 0, 'ieee-be');
-numCols = fread(file, 1, 'int32', 0, 'ieee-be');
-
-images = fread(file, inf, 'unsigned char');
-images = reshape(images, numCols, numRows, numImages);
-images = permute(images,[2 1 3]);
-
-fclose(file);
-
-% Reshape to #pixels x #examples
-images = reshape(images, size(images, 1) * size(images, 2), size(images, 3));
-% Convert to double and rescale to [0,1]
-images = double(images) / 255;
-
-end
-

### 2016-06-06 coproject/matlab/loadLabels.m

@@ -1,16 +1,0 @@
-function labels = loadLabels(fileName)
-% This function is used to load the labels given a file
-
-file = fopen(fileName, 'rb');
-
-magicNumber = fread(file, 1, 'int32', 0, 'ieee-be');
-assert(magicNumber == 2049, ['Bad magic number in ', fileName, '']);
-
-numLabels = fread(file, 1, 'int32', 0, 'ieee-be');
-
-labels = fread(file, inf, 'unsigned char');
-
-fclose(file);
-
-end
-

### 2016-06-06 coproject/matlab/train.m

@@ -1,80 +1,0 @@
-%% LOADING DATA
-clear all;
-close all;
-clc;
-
-trainSet = loadImages('./data/train-images.idx3-ubyte')';
-trainLabel = strtrim(cellstr(num2str(loadLabels('./data/train-labels.idx1-ubyte'))));
-%testSet = loadImages('./data/t10k-images.idx3-ubyte')';
-%testLabel = strtrim(cellstr(num2str(loadLabels('./data/t10k-labels.idx1-ubyte'))));
-
-SVMModelsLin = cell(10,1);
-SVMModelsPol = cell(10,1);
-SVMModelsRBF = cell(10,1);
-classes = unique(trainLabel);
-
-rng(1); % For reproductibility
-
-trainingTimeLin = 0;
-trainingTimePol = 0;
-trainingTimeRBF = 0;
-
-times = zeros(10,3);
-
-%% TRAINING LIN
-
-display('Start training for Linear Kernel');
-
-for i = 1:numel(classes)
-    display(['Start training for digit ', num2str(i-1)]);
-    idx = strcmp(trainLabel, classes(i)); % Create binary classes for each classifier
-    
-    tic;
-    SVMModelsLin{i} = fitcsvm(trainSet , idx , 'KernelFunction', 'linear');
-    times(i,1) = toc;
-    display(['Finished training for digit ', num2str(i-1), '; Time taken: ', num2str(times(i))]);
-    trainingTimeLin = trainingTimeLin + times(i,1);
-end
-
-save('SVMModelsLin', 'SVMModelsLin','-v7.3');
-save('times', 'times');
-
-%% TRAINING POL
-
-display('Start training for Polynomial Kernel');
-
-load('times');
- 
-for i = 1:numel(classes)
-    display(['Start training for digit ', num2str(i-1)]);
-    idx = strcmp(trainLabel, classes(i)); % Create binary classes for each classifier
-  
-    tic;
-    SVMModelsPol{i} = fitcsvm(trainSet , idx , 'KernelFunction', 'polynomial', 'PolynomialOrder', 4);
-    times(i,2) = toc;
-    display(['Finished training for digit ', num2str(i-1), '; Time taken: ', num2str(times(i))]);
-    trainingTimePol = trainingTimePol + times(i,2);
-end
-
-save('SVMModelsPol', 'SVMModelsPol','-v7.3');
-save('times', 'times');
-
-%% TRAINING RBF
-
-display('Start training for RBF Kernel');
-
[…]

### 2016-06-06 coproject/figures/linear.tex
UPD: Added some post processing and written some part of the report

% This file was created by matlab2tikz.
% Minimal pgfplots version: 1.3
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\begin{tikzpicture}

\begin{axis}[%
width=4.585417in,
height=3.591094in,
at={(0.769167in,0.484687in)},
scale only axis,
axis on top,
xmin=0.5,
xmax=10.5,
xtick={1,2,3,4,5,6,7,8,9,10},
xticklabels={{0},{1},{2},{3},{4},{5},{6},{7},{8},{9}},
xlabel={Truth},
y dir=reverse,
ymin=0.5,
ymax=10.5,
ytick={1,2,3,4,5,6,7,8,9,10},
yticklabels={{0},{1},{2},{3},{4},{5},{6},{7},{8},{9}},
ylabel={Prediction}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=10.5,ymin=0.5,ymax=10.5] {linear-1.png};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:1,1,0) {968};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,2,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,3,0) {10};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,4,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,5,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,6,0) {8};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,7,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,8,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,9,0) {3};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,10,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,1,0) {0};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:2,2,0) {1124};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,3,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,4,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,5,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,6,0) {4};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,7,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,8,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,9,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,10,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,1,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,2,0) {2};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:3,3,0) {970};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,4,0) {17};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,5,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,6,0) {4};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,7,0) {9};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,8,0) {21};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,9,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,10,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,1,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,2,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,3,0) {9};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:4,4,0) {952};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,5,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,6,0) {32};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,7,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,8,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,9,0) {23};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,10,0) {13};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,1,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,2,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,3,0) {6};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,4,0) {0};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:5,5,0) {942};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,6,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,7,0) {3};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,8,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,9,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,10,0) {29};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,1,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,2,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,3,0) {3};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,4,0) {14};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,5,0) {0};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:6,6,0) {807};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,7,0) {13};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,8,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,9,0) {28};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,10,0) {3};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,1,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,2,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,3,0) {9};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,4,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,5,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,6,0) {9};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:7,7,0) {923};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,8,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,9,0) {6};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,10,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,1,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,2,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,3,0) {8};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,4,0) {12};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,5,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,6,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,7,0) {0};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:8,8,0) {966};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,9,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,10,0) {15};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,1,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,2,0) {3};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,3,0) {14};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,4,0) {13};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,5,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,6,0) {19};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,7,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,8,0) {1};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:9,9,0) {888};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,10,0) {4};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,1,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,2,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,3,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,4,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,5,0) {23};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,6,0) {3};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,7,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,8,0) {19};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,9,0) {4};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:10,10,0) {931};
\end{axis}
\end{tikzpicture}%

### 2016-06-06 coproject/figures/train-set.tex
UPD: Added some post processing and written some part of the report

% This file was created by matlab2tikz.
% Minimal pgfplots version: 1.3
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\begin{tikzpicture}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(6.367579cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 6}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-1.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(0cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 8}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-2.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(4.245052cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 8}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-3.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(0cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 4}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-4.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(8.490105cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 2}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-5.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(2.122526cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 8}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-6.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(8.490105cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 1}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-7.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(2.122526cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 6}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-8.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(8.490105cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 2}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-9.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(6.367579cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 4}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-10.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(0cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 9}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-11.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(4.245052cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 7}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-12.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(2.122526cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 4}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-13.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(4.245052cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 6}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-14.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(6.367579cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 8}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-15.png};
\end{axis}
\end{tikzpicture}%

### 2016-06-06 coproject/report_2.tex
UPD: Added some post processing and written some part of the report

@@
 \documentclass[a4paper, 11pt]{article}
+% pour l'inclusion de figures en eps,pdf,jpg
+\usepackage{graphicx}
+% quelques symboles mathematiques en plus
+\usepackage{amsmath}
+% le tout en langue francaise
+%\usepackage[french]{babel}
 \usepackage[english]{babel}
+\usepackage[utf8]{inputenc} 
+\usepackage[T1]{fontenc}
+% on utilise les liens internet
+\usepackage{url}
+% pour l'inclusion de links dans le document 
+\usepackage[colorlinks,bookmarks=false,linkcolor=blue,urlcolor=blue]{hyperref}
 
-\usepackage{amsmath}
+\usepackage{tikz}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{pgfplots}
+\usepackage{listings}
+
+\usepackage{float}
+
+\usetikzlibrary{patterns,arrows,decorations.pathreplacing}
+
 \usepackage{amsfonts}
 \usepackage{amssymb}
-\usepackage{xcolor}
-\usepackage{hyperref}
+
+\usepackage{verbatim}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% ATTENTION : package supplementaire
+\usepackage[stable]{footmisc}
+\usepackage{wrapfig}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\paperheight=297mm
+\paperwidth=210mm
+
+\setlength{\textheight}{235mm}
+\setlength{\topmargin}{-1.2cm} % pour centrer la page verticalement
+%\setlength{\footskip}{5mm}
+\setlength{\textwidth}{15cm}
+\setlength{\oddsidemargin}{0.56cm}
+\setlength{\evensidemargin}{0.56cm}
+\setlength{\parindent}{0cm}
 
@@
 
-\title{Kernel Methods and SVMs:
-A convex optimization point of view}
+\title{{\bf Kernel Methods and SVMs} \\
+{\large A convex optimization point of view}}
 \author{Gael Lederrey\\
@@
 impact on further computations involving the best fit.
+
+\subsection{Use case}
+
+The theory works well, but a use case is always needed.
+We decided to use the SVM to learn our computer how to recognize handwritten digits.
+We used the MNIST database.
+Some examples are given in Figure \ref{fig:train-set}.
+
+\begin{figure}[H]
+\centering
+\input{./figures/train-set.tex}
+\caption{\label{fig:train-set} Some examples of the MNIST train set with their label.}
+\end{figure}
+
+We trained the algorithm with different kernel functions:
+\begin{itemize}
+\item Linear kernel: $K(x_i,x_j) = \langle x_i, x_j \rangle$
+\item Polynomial kernel: $K(x_i,x_j) = (\gamma \langle x_i, x_j \rangle + r)^p$, $\gamma$ is a parameter that needs to be found and we decided to keep $r=0$ and we used $p=4$.
+\item RBF kernel:  $K(x_i,x_j) = \exp(-\gamma|x_i-x_j|^2)$, $\gamma$ is a parameter that needs to be found.
+\end{itemize}
+
+For each of these kernels, we trained on train-set composed of 60000 handwritten digits with their labels.
+Then, we test our trained models against a test-set composed of 10000 images.
+We can then compared the predicted labels against the true labels.
+\\\\
+On of the important result is the percentage of correct prediction.
+It is given for each of the kernel in Table \ref{tab:results}.
+As we can see, the best kernel is the RBF one.
+It has already been shown in the literature that this kernel is the best one for SVM.
+\begin{table}[H]
+\centering
+\begin{tabular}{|l|c|}
+\hline
+Linear Kernel & 94.62\% \\\hline
+Polynomial Kernel (degree 4) & 97.31\% \\\hline
+RBF Kernel & 98.56 \% \\ \hline
+\end{tabular}
+\caption{\label{tab:results} Percentage of correct prediction after training.}
+\end{table}
+
+{\color{red} Ici, je veux mettre une figure avec toutes les predictions dur RBF. Et finalement montrer quelques fausses predictions. Juste apres, on pourra faire un conclusion.}
+
 \end{document}

### 2016-06-06 coproject/code/preview.m

@@ -1,0 +1,13 @@
+close all;
+clear all;
+
+images = loadMNISTImages('data/train-images-idx3-ubyte');
+labels = loadMNISTLabels('data/train-labels-idx1-ubyte');
+
+figure()
+rnd = randi(length(labels),15);
+for i=1:15
+    subplot(3,5,i)
+    imshow(reshape(images(:,rnd(i)),[28,28]))
+    title(['label: ', num2str(labels(rnd(i)))])
+end

### 2016-06-06 coproject/code/results.m

@@ -1,0 +1,49 @@
+clear all;
+close all;
+
+linear = [968 0 1 1 1 7 1 1 0 0; 0 1124 2 2 0 1 2 1 3 0; 10 2  970 9 6 3 9 8 14 1; ...
+    1 0 17 952 0 14 0 12 13 1; 1 0 5 0 942 0 7 2 2 23; 8 4 4 32 5 807 9 1 19 3; ...
+    7 2 9 1 3 13 923 0 0 0; 2 5 21 7 7 0 0 966 1 19; 3 5 5 23 7 28 6 5 888 4; ...
+    7 5 1 13 29 3 1 15 4 931];
+
+norm_linear = linear*diag(1./sum(linear',1));
+
+result_linear = sum(diag(norm_linear))/10
+
+figure()
+imagesc(norm_linear);
+colormap(flipud(gray));
+textStrings = num2str(linear(:),'%0.0f');  %# Create strings from the matrix values
+textStrings = strtrim(cellstr(textStrings));  %# Remove any space padding
+[x,y] = meshgrid(1:10);
+hStrings = text(x(:),y(:),textStrings(:),'HorizontalAlignment','center');
+midValue = mean(get(gca,'CLim'));
+textColors = repmat(norm_linear(:) > midValue,1,3);  
+set(hStrings,{'Color'},num2cell(textColors,2));
+
+set(gca,'XTick',1:10,...                         %# Change the axes tick marks
+        'XTickLabel',{'0','1','2','3','4','5','6','7','8','9'},...  %#   and tick labels
+        'YTick',1:10,...
+        'YTickLabel',{'0','1','2','3','4','5','6','7','8','9'},...
+        'TickLength',[0 0]);
+ylabel('Prediction')
+xlabel('Truth')
+
+pol = [970 0 1 0 0 5 1 1 2 0; 0 1123 2 2 1 0 4 0 3 0; 9 6 999 0 2 0 4 8 4 0; ...
+    1 5 3 981 0 6 0 6 5 3; 3 2 3 0 960 0 4 1 1 8; 3 2 2 13 1 859 4 1 5 2; ...
+    4 6 2 0 4 6 934 0 2 0; 0 14 8 2 1 0 0 997 0 6; 7 2 1 6 5 6 1 2 942 2; ...
[…]

### 2016-06-07 coproject/figures/wrong-label.tex
UPD: Finished writing experience + conclusion. Just need to review the whole report

% This file was created by matlab2tikz.
% Minimal pgfplots version: 1.3
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\begin{tikzpicture}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(8.490105cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={4 $\rightarrow$ 2}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-1.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(0cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={5 $\rightarrow$ 3}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-2.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(6.367579cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={9 $\rightarrow$ 3}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-3.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(6.367579cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={3 $\rightarrow$ 7}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-4.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(0cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={2 $\rightarrow$ 0}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-5.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(8.490105cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={7 $\rightarrow$ 9}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-6.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(8.490105cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={9 $\rightarrow$ 3}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-7.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(4.245052cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={5 $\rightarrow$ 3}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-8.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(2.122526cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={9 $\rightarrow$ 7}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-9.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(4.245052cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={4 $\rightarrow$ 6}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-10.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(6.367579cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={7 $\rightarrow$ 2}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-11.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(0cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={2 $\rightarrow$ 7}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-12.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(4.245052cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={4 $\rightarrow$ 2}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-13.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(2.122526cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={9 $\rightarrow$ 3}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-14.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(2.122526cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={3 $\rightarrow$ 7}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-15.png};
\end{axis}
\end{tikzpicture}%

### 2016-06-07 coproject/report_tallec_lederrey.tex
UPD: Finished writing experience + conclusion. Just need to review the whole report

--- 2016-06-06 coproject/report_2.tex
@@
 \usepackage{tikz}
+\usepackage{footnote}
 \usepackage{caption}
@@
 \usepackage{listings}
+\usepackage{array,hhline}
+\usepackage{pgf}
+\usepackage{multirow,graphicx,array,xcolor,colortbl}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\newcommand{\cca}[2]{
+	\cellcolor{black!#1}
+	\ifnum #1>50
+		\color{white}
+	\fi
+	{#2}
+	}
 
@@
   machines special case.
-
-  Most theoretical results used here are presented, explained and
-  proved more thoroughly in the ``Master vision et apprentissage''
-  of the ``Ecole normale superieure de Cachan'' course on kernels.
 \end{abstract}
 
-\section{Kernel methods}
+\section{Kernel methods\footnotemark}
 \subsection{Generalities}
 \label{sub:gen}
+
+\footnotetext{The different elements presented in this section comes from the course on Machine Learning \cite{course} as well as the paper explaining how convex optimization is used with Kernel methods\cite{paper}.}
 
@@
 
-\section{Support vector machines: a special case}
+\section{Support vector machines: a special case\footnotemark}
+\footnotetext{The different elements presented in this section comes from the course on Machine Learning \cite{course} as well as the article on Wikip\'edia on SVM\cite{svm}.}
 Support vector machines are kernel methods used in the field of binary
@@
 We decided to use the SVM to learn our computer how to recognize handwritten digits.
-We used the MNIST database.
-Some examples are given in Figure \ref{fig:train-set}.
+To do so, we create a script in python using the {\it sklearn}\cite{python} library.
+We used the MNIST database of handwritten digits\cite{mnist} provided by Yann Lecun and well studied.
+Each of the images (a handwritten digit between 0 and 9) has a given label.
+The purpose is to create a model using SVM such that the computer can recognize these digits.
+Some examples of the MNIST dataset are given in Figure \ref{fig:train-set}.
 
@@
 
-We trained the algorithm with different kernel functions:
+The purpose of this script was to train the computer on 60000 images with their labels.
+And then, test the model on 10000 images.
+We could finally compare the predicted labels against the real labels.
+In order to test the SVM method, we trained the algorithm with different kernel functions:
 \begin{itemize}
@@
 
-For each of these kernels, we trained on train-set composed of 60000 handwritten digits with their labels.
-Then, we test our trained models against a test-set composed of 10000 images.
-We can then compared the predicted labels against the true labels.
-\\\\
-On of the important result is the percentage of correct prediction.
-It is given for each of the kernel in Table \ref{tab:results}.
-As we can see, the best kernel is the RBF one.
-It has already been shown in the literature that this kernel is the best one for SVM.
+The most important result is the correctness of the prediction.
+To do this, we take the number of correct prediction and we divide it by the length of the dataset.
+This gives us a percentage on the correctness of the model.
+We give in Table \ref{tab:correct} the results for each of the kernels.
+
 \begin{table}[H]
 \centering
-\begin{tabular}{|l|c|}
+\begin{tabular}{l|c}
+Kernels & Correct prediction \\
 \hline
-Linear Kernel & 94.62\% \\\hline
-Polynomial Kernel (degree 4) & 97.31\% \\\hline
-RBF Kernel & 98.56 \% \\ \hline
+Linear & 94.62\% \\
+Polynomial (degree 4) & 97.31\% \\ 
+RBF & 98.56 \% \\ 
 \end{tabular}
-\caption{\label{tab:results} Percentage of correct prediction after training.}
+\caption{\label{tab:correct} Percentage of correct prediction for each of the kernels.}
 \end{table}
 
-{\color{red} Ici, je veux mettre une figure avec toutes les predictions dur RBF. Et finalement montrer quelques fausses predictions. Juste apres, on pourra faire un conclusion.}
+As we can see, the best kernel is the RBF one.
+It has already been shown in the literature that this kernel is the state-of-the-art in SVM.
+Another important aspect is to have a look at the predictions more precisely.
+For this, we use a confusion matrix.
+This matrix shows the labels (vertical axis) in function of the predictions (horizontal axis).
+It is interesting to have a look at this matrix to see if there is some bias.
+(For example, 1 can be often predicted as 7).
+The confusion matrix for the RBF kernel is given in Table \ref{tab:conf_mat}.
+As we can see in this table, we don't have any bias.
+We just show some wrong predictions in Figure \ref{fig:wrong-label}.
+This figure shows us that some of the images are difficult to label, even for us.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{cc|c|c|c|c|c|c|c|c|c|c|}
+\multicolumn{2}{c}{} & \multicolumn{10}{c}{Predictions} \\
+\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{3} & \multicolumn{1}{c}{4} & \multicolumn{1}{c}{5} & \multicolumn{1}{c}{6} & \multicolumn{1}{c}{7} & \multicolumn{1}{c}{8} & \multicolumn{1}{c}{9}  \\
+\cline{3-12}
+\multirow{10}{*}{\rotatebox[origin=c]{90}{Labels}}
+& 0 & \cca{99}{974} & 0 & 1 & 0 & 0 & 2 & 0 & 1 & 2 & 0 \\ \cline{3-12}
+& 1 & 0 & \cca{99}{1128} & 3 & 1 & 0 & 1 & 0 & 1 & 1 & 0 \\ \cline{3-12}
+& 2 & 4 & 0 & \cca{98}{1017} & 0 & 1 & 0 & 0 & 7 & 3 & 0 \\ \cline{3-12}
+& 3 & 0 & 0 & 2 & \cca{99}{997} & 1 & 2 & 0 & 4 & 3 & 1 \\ \cline{3-12}
+& 4 & 0 & 0 & 2 & 0 & \cca{99}{968} & 0 & 4 & 0 & 1 & 7 \\ \cline{3-12} 
+& 5 & 2 & 0 & 0 & 5 & 1 & \cca{98}{877} & 3 & 1 & 2 & 1 \\ \cline{3-12} 
+& 6 & 3 & 2 & 0 & 0 & 2 & 2 & \cca{99}{948} & 0 & 1 & 0 \\ \cline{3-12} 
+& 7 & 0 & 3 & 8 & 1 & 1 & 0 & 0 & \cca{98}{1007} & 1 & 7 \\ \cline{3-12} 
+& 8 & 3 & 0 & 1 & 3 & 1 & 1 & 0 & 2 & \cca{97}{959} & 4 \\ \cline{3-12} 
+& 9 & 2 & 3 & 1 & 6 & 6 & 2 & 1 & 5 & 1 & \cca{100}{982} \\ \cline{3-12} 
+\end{tabular}
+\caption{\label{tab:conf_mat} Confusion Matrix for the RBF kernel.}
+\end{table}
+
+\begin{figure}[H]
+\centering
+\input{./figures/wrong-label.tex}
+\caption{\label{fig:wrong-label} Examples of images that have not been labelled correctly. This number above the pictures are like this: label $\rightarrow$ prediction.}
+\end{figure}
+
+\section{Conclusion}
+
+Within this project, we explored a really interesting topic.
+Indeed, Machine Learning as well as Artificial Intelligence is today's future technology.
+A lot of start-ups are working on this and applying it to a lot of different fields such as Facial and Emotion Recognition, Language Recognition, Data Mining, etc.
+Until not so long ago, the SVM were the state-of-the-art.
+Now, Random Forests and Deep-Learning (Convolutional Neural Networks) are better than this method.
+But their underlying theory is sometimes not totally understood compared to the SVM.
+Indeed, Convex Optimization is very known field by many scientist.
+Developing a model using Convex Optimization to make the computer learn things, such as handwritten digits, is quite amazing.
+
+\begin{thebibliography}{1}
+
+\bibitem{course} \href{http://lear.inrialpes.fr/people/mairal/teaching/2015-2016/MVA/fichiers/mva_slides.pdf}{Slides of the course on "{\it Kernel Methods for Machine Learning}", Inria, Grenoble, France}
+
+\bibitem{paper} Kim, Seung-Jean, et al. "Learning the kernel via convex optimization." {\it Acoustics, Speech and Signal Processing, 2008. ICASSP 2008. IEEE International Conference on.} IEEE, 2008.
+
+\bibitem{svm} \href{https://en.wikipedia.org/wiki/Support_vector_machine}{Wikip\'edia - Support Vector Machine}
+
+\bibitem{python} \href{Library {\it sklearn} for Machine Learning in Python. Functions on SVM/SVC.}{http://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html}
+
+\bibitem{mnist} \href{The MNIST Databse of handwritten digits. (Website of Yann Lecun)}{http://yann.lecun.com/exdb/mnist/}
+
+\end{thebibliography}
 

### 2016-06-07 coproject/code/results.m

@@ -10,25 +10,6 @@ norm_linear = linear*diag(1./sum(linear',1));
 
 result_linear = sum(diag(norm_linear))/10
 
-figure()
-imagesc(norm_linear);
-colormap(flipud(gray));
-textStrings = num2str(linear(:),'%0.0f');  %# Create strings from the matrix values
-textStrings = strtrim(cellstr(textStrings));  %# Remove any space padding
-[x,y] = meshgrid(1:10);
-hStrings = text(x(:),y(:),textStrings(:),'HorizontalAlignment','center');
-midValue = mean(get(gca,'CLim'));
-textColors = repmat(norm_linear(:) > midValue,1,3);  
-set(hStrings,{'Color'},num2cell(textColors,2));
-
-set(gca,'XTick',1:10,...                         %# Change the axes tick marks
-        'XTickLabel',{'0','1','2','3','4','5','6','7','8','9'},...  %#   and tick labels
-        'YTick',1:10,...
-        'YTickLabel',{'0','1','2','3','4','5','6','7','8','9'},...
-        'TickLength',[0 0]);
-ylabel('Prediction')
-xlabel('Truth')
-
 pol = [970 0 1 0 0 5 1 1 2 0; 0 1123 2 2 1 0 4 0 3 0; 9 6 999 0 2 0 4 8 4 0; ...
     1 5 3 981 0 6 0 6 5 3; 3 2 3 0 960 0 4 1 1 8; 3 2 2 13 1 859 4 1 5 2; ...
     4 6 2 0 4 6 934 0 2 0; 0 14 8 2 1 0 0 997 0 6; 7 2 1 6 5 6 1 2 942 2; ...
@@ -47,3 +28,22 @@ rbf = [974 0 1 0 0 2 0 1 2 0; 0 1128 3 1 0 1 0 1 1 0; 4 0 1017 0 1 0 0 7 3 0; ..
 norm_rbf = rbf*diag(1./sum(rbf',1));
 
 result_rbf = sum(diag(norm_rbf))/10
+
+figure()
+imagesc(norm_rbf);
+colormap(flipud(gray));
+textStrings = num2str(rbf(:),'%0.0f');  %# Create strings from the matrix values
+textStrings = strtrim(cellstr(textStrings));  %# Remove any space padding
+[x,y] = meshgrid(1:10);
+hStrings = text(x(:),y(:),textStrings(:),'HorizontalAlignment','center');
+midValue = mean(get(gca,'CLim'));
[…]

### 2016-06-07 coproject/code/wrong.m

@@ -1,0 +1,16 @@
+close all;
+clear all;
+
+idx = [247, 321, 340, 381, 445, 448, 495, 582, 684, 720, 740, 882, 947, 951, 956, 1014, 1039, 1044, 1112, 1226, 1232, 1242, 1247, 1299, 1319, 1328, 1378, 1393, 1500, 1522, 1530, 1549, 1553, 1581, 1681, 1709, 1754, 1790, 1901, 2035, 2044, 2070, 2098, 2109, 2118, 2129, 2130, 2135, 2182, 2185, 2272, 2293, 2371, 2387, 2422, 2488, 2597, 2607, 2648, 2654, 2863, 2896, 2927, 2939, 2953, 3060, 3073, 3117, 3289, 3333, 3422, 3475, 3503, 3520, 3558, 3559, 3597, 3767, 3776, 3780, 3796, 3808, 3853, 3906, 3941, 3976, 4065, 4075, 4078, 4163, 4176, 4199, 4201, 4248, 4289, 4443, 4497, 4500, 4639, 4740, 4807, 4823, 4880, 4966, 5457, 5642, 5734, 5887, 5936, 5937, 5955, 5973, 6035, 6059, 6166, 6555, 6560, 6571, 6576, 6597, 6625, 6651, 7434, 8094, 8527, 9009, 9015, 9024, 9587, 9634, 9642, 9664, 9692, 9729, 9745, 9749, 9768, 9770, 9779, 9792, 9808, 9839, 9944];
[…]

### 2016-06-10 coproject/report_tallec_lederrey.tex
Final changes to the report

@@
 
-The theory works well, but a use case is always needed.
-We decided to use the SVM to learn our computer how to recognize handwritten digits.
-To do so, we create a script in python using the {\it sklearn}\cite{python} library.
-We used the MNIST database of handwritten digits\cite{mnist} provided by Yann Lecun and well studied.
-Each of the images (a handwritten digit between 0 and 9) has a given label.
-The purpose is to create a model using SVM such that the computer can recognize these digits.
-Some examples of the MNIST dataset are given in Figure \ref{fig:train-set}.
+Let us now dwelve in a commonly presented use case. We decided to use the SVM
+to perform handwritten digits recognition handwritten digits.  The {\it
+sklearn}\cite{python} machine learning library was used, to train SVMs on the
+MNIST database of handwritten digits\cite{mnist} retrieved from Yann Lecun's
+website.  Each of the images (a handwritten digit between 0 and 9) has a given
+label. The goal is to train a SVM to  recognize these digits. Some examples of
+the MNIST dataset are given in
+Figure \ref{fig:train-set}.
 
@@
 
-The purpose of this script was to train the computer on 60000 images with their labels.
-And then, test the model on 10000 images.
-We could finally compare the predicted labels against the real labels.
-In order to test the SVM method, we trained the algorithm with different kernel functions:
+The purpose of this script was to train SVMs on 60000 images with their
+labels. And then, test the model on 10000 images. We could finally compare the
+predicted labels against the real labels. In order to test the SVM method, we
+trained the algorithm with different kernel functions. As SVMs naturally perform
+binary classification, and not $n$-ary classification, one-vs-all classification
+was used (that is we train the model n-times, once on each digit).
 \begin{itemize}
@@
 
-The most important result is the correctness of the prediction.
-To do this, we take the number of correct prediction and we divide it by the length of the dataset.
-This gives us a percentage on the correctness of the model.
-We give in Table \ref{tab:correct} the results for each of the kernels.
+The most important result is the correctness of the prediction. To do this, we
+take the number of correct prediction and we divide it by the length of the
+dataset. This gives us a percentage on the correctness of the model. We give in
+Table \ref{tab:correct} the results for each of the kernels.
 
@@
 
-As we can see, the best kernel is the RBF one.
-It has already been shown in the literature that this kernel is the state-of-the-art in SVM.
-Another important aspect is to have a look at the predictions more precisely.
-For this, we use a confusion matrix.
-This matrix shows the labels (vertical axis) in function of the predictions (horizontal axis).
-It is interesting to have a look at this matrix to see if there is some bias.
-(For example, 1 can be often predicted as 7).
-The confusion matrix for the RBF kernel is given in Table \ref{tab:conf_mat}.
-As we can see in this table, we don't have any bias.
-We just show some wrong predictions in Figure \ref{fig:wrong-label}.
-This figure shows us that some of the images are difficult to label, even for us.
+As we can see, the best kernel is the RBF one. It has already been shown in the
+literature that this kernel is the state-of-the-art in SVM. Another important
+aspect is to have a look at the predictions more precisely. For this, we use a
+confusion matrix. This matrix shows the labels (vertical axis) in function of
+the predictions (horizontal axis). It is interesting to have a look at this
+matrix to see if there is some bias. (For example, 1 can be often predicted as
+7). The confusion matrix for the RBF kernel is given in Table
+\ref{tab:conf_mat}. As we can see in this table, we don't have any bias. We
+just show some wrong predictions in Figure \ref{fig:wrong-label}. This figure
+shows us that some of the images are difficult to label, even for us.
 
@@
 
-Within this project, we explored a really interesting topic.
-Indeed, Machine Learning as well as Artificial Intelligence is today's future technology.
-A lot of start-ups are working on this and applying it to a lot of different fields such as Facial and Emotion Recognition, Language Recognition, Data Mining, etc.
-Until not so long ago, the SVM were the state-of-the-art.
-Now, Random Forests and Deep-Learning (Convolutional Neural Networks) are better than this method.
-But their underlying theory is sometimes not totally understood compared to the SVM.
-Indeed, Convex Optimization is very known field by many scientist.
-Developing a model using Convex Optimization to make the computer learn things, such as handwritten digits, is quite amazing.
+Kernel methods remain methods of interest, most notably in machine learning,
+since they provide both a satisfactory theoretical background, as well as
+some good results in a large variety of fields. One thing that might be worth
+noticing is that they have a quadratic complexity in the number of data-entries
+in the train dataset (we need to be able to compute the whole kernel, which is
+of dimension $n \times n$ where $n$ is the number of train datapoints), making
+them hardly applicable for very large datasets.
+
+This might partly explain the recent arising of deep learning methods, which
+are much less theoretically grounded, and among other ill-understood properties,
+perform well while not achieving global optimality, but keep a linear complexity
+in the number of datapoints, both at train and at test time (even if it might argued
+that the train time probably doesn't scale linearly in the number of training samples).
+
+Kernel methods remain useful baselines when considering small datasets, as
+they remain computationnally tractable, and furthermore
+provide a natural regularization, that tends to reduce efficiently overfitting.
+Besides, neural networks tend to perform quite poorly in those cases where only 
+few datapoints are available as they tend to overfit rapidly, even though recent
+researches tend to bridge the gap.
 

### 2016-06-10 coproject/report_tallec_lederrey.tex
UPD: Report compiled

@@
 
-\setlength{\textheight}{235mm}
-\setlength{\topmargin}{-1.2cm} % pour centrer la page verticalement
+\setlength{\textheight}{240mm}
+\setlength{\topmargin}{-1.5cm} % pour centrer la page verticalement
 %\setlength{\footskip}{5mm}
@@
 was used (that is we train the model n-times, once on each digit).
+\newpage
 \begin{itemize}
@@
 
-\bibitem{paper} Kim, Seung-Jean, et al. "Learning the kernel via convex optimization." {\it Acoustics, Speech and Signal Processing, 2008. ICASSP 2008. IEEE International Conference on.} IEEE, 2008.
+\bibitem{paper} Kim, Seung-Jean, et al. "Learning the kernel via convex optimization." {\it Acoustics, Speech and Signal Processing, ICASSP.} IEEE, 2008.
 
